---
title: "Association between nature exposure and brain health: cross-sectional results
  from the UK Biobank Study"
author: "Ryan Stanley Falck; Revised by Matt Noseworthy"
date: "Revision 7: 2025-06-07"
output: 
    html_document:
      toc: true
      theme: united
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# 1) Initial set-up

Load packages for the analysis using the "pacman" package
```{r}
if (!require("pacman")) install.packages("pacman")
pacman::p_load(boot,MASS,readxl,plyr,ggplot2,lme4,lmerTest,lsmeans,gtsummary,ggthemes,patchwork,modelsummary,psych,
               lavaan,semTools,semPlot,expss)
```

Set the working directory (i.e., where the data are stored)
```{r}
setwd("~/Desktop/OneDrive - UBC/UK Biobank Greenspace Analysis")
```

Load the data set 
* Notes: 
+ 1) This dataset has been reduced from the full UKB dataset to the primary variables of interest for this analysis.
```{r}
fulldata<-read.csv("UKB_Fulldata_Sept27_2023.csv") #Data#
```


# 2) Set-up variables for analysis

Change sex to male vs. female
```{r}
fulldata$Sex[fulldata$Sex==1]<-"Male"
fulldata$Sex[fulldata$Sex==0]<-"Female"
```

Assessment centre
```{r}
fulldata$Assessment_Centre<-NA
fulldata$Assessment_Centre[fulldata$Assessment.centre==11012]<-"Barts"
fulldata$Assessment_Centre[fulldata$Assessment.centre==11021]<-"Birmingham"
fulldata$Assessment_Centre[fulldata$Assessment.centre==11011]<-"Bristol"
fulldata$Assessment_Centre[fulldata$Assessment.centre==11008]<-"Bury"
fulldata$Assessment_Centre[fulldata$Assessment.centre==11003]<-"Cardiff"
fulldata$Assessment_Centre[fulldata$Assessment.centre==11024]<-"Cheadle (revisit)"
fulldata$Assessment_Centre[fulldata$Assessment.centre==11020]<-"Croydon"
fulldata$Assessment_Centre[fulldata$Assessment.centre==11005]<-"Edinburgh"
fulldata$Assessment_Centre[fulldata$Assessment.centre==11004]<-"Glasgow"
fulldata$Assessment_Centre[fulldata$Assessment.centre==11018]<-"Hounslow"
fulldata$Assessment_Centre[fulldata$Assessment.centre==11010]<-"Leeds"
fulldata$Assessment_Centre[fulldata$Assessment.centre==11016]<-"Liverpool"
fulldata$Assessment_Centre[fulldata$Assessment.centre==11001]<-"Manchester"
fulldata$Assessment_Centre[fulldata$Assessment.centre==11017]<-"Middlesborough"
fulldata$Assessment_Centre[fulldata$Assessment.centre==11009]<-"Newcastle"
fulldata$Assessment_Centre[fulldata$Assessment.centre==11013]<-"Nottingham"
fulldata$Assessment_Centre[fulldata$Assessment.centre==11002]<-"Oxford"
fulldata$Assessment_Centre[fulldata$Assessment.centre==11007]<-"Reading"
fulldata$Assessment_Centre[fulldata$Assessment.centre==11014]<-"Sheffield"
fulldata$Assessment_Centre[fulldata$Assessment.centre==10003]<-"Stockport (pilot)"
fulldata$Assessment_Centre[fulldata$Assessment.centre==11006]<-"Stoke"
fulldata$Assessment_Centre[fulldata$Assessment.centre==11022]<-"Swansea"
fulldata$Assessment_Centre[fulldata$Assessment.centre==11023]<-"Wrexham"
fulldata$Assessment_Centre[fulldata$Assessment.centre==11025]<-"Cheadle (imaging)"
fulldata$Assessment_Centre[fulldata$Assessment.centre==11026]<-"Reading (imaging)"
fulldata$Assessment_Centre[fulldata$Assessment.centre==11027]<-"Newcastle (imaging)"
fulldata$Assessment_Centre[fulldata$Assessment.centre==11028]<-"Bristol (imaging)"
```

Length of time at current address (3 timepoints [T0, T2, T3])
* Recorded in whole years. -10: less than a year; -1: do not know; -3: prefer not to answer
```{r}
fulldata$Time_at_current_address_T0 <- fulldata$Length.of.time.at.current.address.0
fulldata$Time_at_current_address_T0[fulldata$Time_at_current_address_T0==-10] <- 0.5
fulldata$Time_at_current_address_T0[fulldata$Time_at_current_address_T0==-1 | fulldata$Time_at_current_address_T0==-3] <- NA

fulldata$Time_at_current_address_T2 <- fulldata$Length.of.time.at.current.address.2
fulldata$Time_at_current_address_T2[fulldata$Time_at_current_address_T2==-10] <- 0.5
fulldata$Time_at_current_address_T2[fulldata$Time_at_current_address_T2==-1 | fulldata$Time_at_current_address_T2==-3] <- NA

fulldata$Time_at_current_address_T3 <- fulldata$Length.of.time.at.current.address.3
fulldata$Time_at_current_address_T3[fulldata$Time_at_current_address_T3==-10] <- 0.5
fulldata$Time_at_current_address_T3[fulldata$Time_at_current_address_T3==-1 | fulldata$Time_at_current_address_T3==-3] <- NA
```

Household Income
```{r}
fulldata$Income<-NA
fulldata$Income[fulldata$Household.Income==1]<- "<18,000 GBP"
fulldata$Income[fulldata$Household.Income==2]<- "18,000-30,999 GBP"
fulldata$Income[fulldata$Household.Income==3]<- "31,000-51,999 GBP"
fulldata$Income[fulldata$Household.Income==4]<- "52,000-100,000 GBP"
fulldata$Income[fulldata$Household.Income==5]<- ">100,000 GBP"
fulldata$Income[fulldata$Household.Income<=0]<- "Do not know or prefer not to answer"
```


Time spent outdoors (categorical variable)
```{r}
#Time spent outdoors in summer (categorical)
fulldata$Time_outdoors_summer<-NA
fulldata$Time_outdoors_summer[fulldata$Time.spent.outdoors.in.summer==0 | fulldata$Time.spent.outdoors.in.summer==-10]<-"Less than 1 hour/day"
fulldata$Time_outdoors_summer[fulldata$Time.spent.outdoors.in.summer==-3 | fulldata$Time.spent.outdoors.in.summer==-1]<-"Do not know or prefer not to say"
fulldata$Time_outdoors_summer[fulldata$Time.spent.outdoors.in.summer>=1 & fulldata$Time.spent.outdoors.in.summer<7]<-"1-6 hours/day"
fulldata$Time_outdoors_summer[fulldata$Time.spent.outdoors.in.summer>=7 & fulldata$Time.spent.outdoors.in.summer<13]<-"7-12 hours/day"
fulldata$Time_outdoors_summer[fulldata$Time.spent.outdoors.in.summer>=13]<-"13+ hours/day"

#Time spent outdoors in winter (categorical)
fulldata$Time_outdoors_winter<-NA
fulldata$Time_outdoors_winter[fulldata$Time.spent.outdoors.in.winter==0 | fulldata$Time.spent.outdoors.in.winter==-10]<-"Less than 1 hour/day"
fulldata$Time_outdoors_winter[fulldata$Time.spent.outdoors.in.winter==-3 | fulldata$Time.spent.outdoors.in.winter==-1]<-"Do not know or prefer not to say"
fulldata$Time_outdoors_winter[fulldata$Time.spent.outdoors.in.winter>=1 & fulldata$Time.spent.outdoors.in.winter<7]<-"1-6 hours/day"
fulldata$Time_outdoors_winter[fulldata$Time.spent.outdoors.in.winter>=7 & fulldata$Time.spent.outdoors.in.winter<13]<-"7-12 hours/day"
fulldata$Time_outdoors_winter[fulldata$Time.spent.outdoors.in.winter>=13]<-"13+ hours/day"
```


Time spent outdoors (continuous variable)
```{r}
#Time spent outdoors in summer (continuous)
fulldata$Time_outdoors_summer_cont<-fulldata$Time.spent.outdoors.in.summer
fulldata$Time_outdoors_summer_cont[fulldata$Time_outdoors_summer_cont==-10]<-0.5
fulldata$Time_outdoors_summer_cont[fulldata$Time_outdoors_summer_cont==-1 | fulldata$Time_outdoors_summer_cont==-3]<-NA

#Time spent outdoors in winter (continuous)
fulldata$Time_outdoors_winter_cont<-fulldata$Time.spent.outdoors.in.winter
fulldata$Time_outdoors_winter_cont[fulldata$Time_outdoors_winter_cont==-10]<-0.5
fulldata$Time_outdoors_winter_cont[fulldata$Time_outdoors_winter_cont==-1 | fulldata$Time_outdoors_winter_cont==-3]<-NA

#Time spent outdoors, averaged over summer and winter (continuous)
fulldata$Average_time_outdoors <- (fulldata$Time_outdoors_summer_cont + fulldata$Time_outdoors_winter_cont) / 2
```


Disability or infirmity status
```{r}
fulldata$Disability<-NA
fulldata$Disability[fulldata$Disability.or.Infirmity.status==1]<-"Yes"
fulldata$Disability[fulldata$Disability.or.Infirmity.status==0]<-"No"
fulldata$Disability[fulldata$Disability.or.Infirmity.status<=-1]<-"Don't know or prefer not to answer"
```


Hearing difficulty status
```{r}
fulldata$Hearing_difficulty<-NA
fulldata$Hearing_difficulty[fulldata$Hearing.Difficulty==99]<-"Completely Deaf"
fulldata$Hearing_difficulty[fulldata$Hearing.Difficulty==1]<-"Difficulty Hearing"
fulldata$Hearing_difficulty[fulldata$Hearing.Difficulty==0]<-"No Difficulty"
fulldata$Hearing_difficulty[fulldata$Hearing.Difficulty<=-1]<-"Don't know or prefer not to answer"
```


Diabetes status
```{r}
fulldata$Diabetes<-NA
fulldata$Diabetes[fulldata$Diabetes.status==1]<-"Yes"
fulldata$Diabetes[fulldata$Diabetes.status==0]<-"No"
fulldata$Diabetes[fulldata$Diabetes.status<=-1]<-"Don't know or prefer not to answer"
```


Cancer status
```{r}
fulldata$Cancer<-NA
fulldata$Cancer[fulldata$Cancer.status==1]<-"Yes"
fulldata$Cancer[fulldata$Cancer.status==0]<-"No"
fulldata$Cancer[fulldata$Cancer.status<=-1]<-"Don't know or prefer not to answer"
```


Other serious medical condition status
```{r}
fulldata$Other_medical_conditions<-NA
fulldata$Other_medical_conditions[fulldata$Other.serious.medical.condition==1]<-"Yes"
fulldata$Other_medical_conditions[fulldata$Other.serious.medical.condition==0]<-"No"
fulldata$Other_medical_conditions[fulldata$Other.serious.medical.condition<=-1]<-"Don't know or prefer not to answer"
```


Educational attainment
```{r}
fulldata$Education[fulldata$Education==1]<-"College or University Degree"
fulldata$Education[fulldata$Education==2]<-"A levels/AS levels or equivalent"
fulldata$Education[fulldata$Education==3]<-"O levels/General Certificate of Secondary Education or equivalent"
fulldata$Education[fulldata$Education==4]<-"CSEs or equivalent"
fulldata$Education[fulldata$Education==5]<-"National Vocational Qualification or equivalent"
fulldata$Education[fulldata$Education==6]<-"Other professional qualifications"
fulldata$Education[fulldata$Education==-7]<-"None of the above"
fulldata$Education[fulldata$Education==-3]<-"Prefer not to answer"
```


Employment status
```{r}
fulldata$Employment[fulldata$Employment==1]<-"In paid employment or self-employed"
fulldata$Employment[fulldata$Employment==2]<-"Retired"
fulldata$Employment[fulldata$Employment==3]<-"Looking after home and/or family"
fulldata$Employment[fulldata$Employment==4]<-"Unable to work because of sickness or disability"
fulldata$Employment[fulldata$Employment==5]<-"Unemployed"
fulldata$Employment[fulldata$Employment==6]<-"Doing unpaid or voluntary work"
fulldata$Employment[fulldata$Employment==7]<-"Student"
fulldata$Employment[fulldata$Employment==-3]<-"Prefer not to answer"
fulldata$Employment[fulldata$Employment==-7]<-"None of the above"
```


Eye problems (data only available from ~30,000 participants)
```{r}
fulldata$Eye_problems<-NA
fulldata$Eye_problems[fulldata$Eye.problems==1 | fulldata$Eye.problems==2 | fulldata$Eye.problems==3 | fulldata$Eye.problems==4 | fulldata$Eye.problems==5 | fulldata$Eye.problems==6]<-"Yes"
fulldata$Eye_problems[fulldata$Eye.problems==-7]<-"No"
fulldata$Eye_problems[fulldata$Eye.problems==-3]<-"Prefer not to answer"
fulldata$Eye_problems[fulldata$Eye.problems==-1]<-"Do not know"
```


Cardiovascular disease status
```{r}
#Cardiovascular Disease
fulldata$CVD_problems<-NA
fulldata$CVD_problems[fulldata$CVD.problems==1 |fulldata$CVD.problems==2 | fulldata$CVD.problems==3 |
                     fulldata$CVD.problems==4]<-"Yes"
fulldata$CVD_problems[fulldata$CVD.problems==-7]<-"No"
fulldata$CVD_problems[fulldata$CVD.problems==-3]<-"Prefer not to answer"
```


Blood clots, allergies, and breathing problems status
```{r}
fulldata$Clots_Asthma_Allergies<-NA
fulldata$Clots_Asthma_Allergies[fulldata$Blood.clots.and.breathing.problems==5 | fulldata$Blood.clots.and.breathing.problems==7 |
                               fulldata$Blood.clots.and.breathing.problems==6 |fulldata$Blood.clots.and.breathing.problems==8 |
                               fulldata$Blood.clots.and.breathing.problems==9]<-"Yes"
fulldata$Clots_Asthma_Allergies[fulldata$Blood.clots.and.breathing.problems==-7] <- "No"
fulldata$Clots_Asthma_Allergies[fulldata$Blood.clots.and.breathing.problems==-3] <- "Prefer not to answer"
```


Alcohol consumption
```{r}
fulldata$Alcohol<-NA
fulldata$Alcohol[fulldata$Alcohol.intake==0]<-"Never"
fulldata$Alcohol[fulldata$Alcohol.intake==1]<-"Previous drinker"
fulldata$Alcohol[fulldata$Alcohol.intake==2]<-"Current drinker"
fulldata$Alcohol[fulldata$Alcohol.intake==-3]<-"Prefer not to answer"
```


Smoking status
```{r}
fulldata$Smoking<-NA
fulldata$Smoking[fulldata$Smoking.Status==0]<-"Never"
fulldata$Smoking[fulldata$Smoking.Status==1]<-"Previous Smoker"
fulldata$Smoking[fulldata$Smoking.Status==2]<-"Current Smoker"
fulldata$Smoking[fulldata$Smoking.Status==-3]<-"Prefer not to answer"
```


Home area population density
```{r}
fulldata$Population_density<-NA
fulldata$Population_density[fulldata$Home.area.population.density==1 | fulldata$Home.area.population.density==5 | 
                fulldata$Home.area.population.density==11 | fulldata$Home.area.population.density==12]<-"Urban area"
fulldata$Population_density[fulldata$Home.area.population.density==2 | fulldata$Home.area.population.density==6 | 
                           fulldata$Home.area.population.density==13]<-"Town"
fulldata$Population_density[fulldata$Home.area.population.density==3 | fulldata$Home.area.population.density==4 | 
                           fulldata$Home.area.population.density==7 | fulldata$Home.area.population.density==8 |
                           fulldata$Home.area.population.density>=14]<-"Village, Hamlet, or Rural Area"
fulldata$Population_density[fulldata$Home.area.population.density==9]<-"Postcode not linkable" 
```


Bipolar/Depression Status
```{r}
fulldata$Bipolar_Depression<-NA
fulldata$Bipolar_Depression[fulldata$Bipolar.and.depression.status==0]<-"No Bipolar or Depression"
fulldata$Bipolar_Depression[fulldata$Bipolar.and.depression.status==1 | 
                           fulldata$Bipolar.and.depression.status==2]<-"Bipolar Disorder"
fulldata$Bipolar_Depression[fulldata$Bipolar.and.depression.status==3 | 
                           fulldata$Bipolar.and.depression.status==4]<-"Probable Recurrent Major Depressive Disorder"
fulldata$Bipolar_Depression[fulldata$Bipolar.and.depression.status==5]<-"Probable Single Major Depressive Episode"
```


Ethnicity/Race
```{r}
fulldata$Race<-NA
fulldata$Race[fulldata$Ethnicity==1001 | fulldata$Ethnicity==1002 | fulldata$Ethnicity==1003]<-"White"
fulldata$Race[fulldata$Ethnicity==2001 | fulldata$Ethnicity==2002 | fulldata$Ethnicity==2003 | fulldata$Ethnicity==2004]<-"Mixed"
fulldata$Race[fulldata$Ethnicity==3001 | fulldata$Ethnicity==3002 | fulldata$Ethnicity==3003 | fulldata$Ethnicity==3004]<-"Asian or Asian British"
fulldata$Race[fulldata$Ethnicity==4001 | fulldata$Ethnicity==4002 | fulldata$Ethnicity==4003]<-"Black or Black British"
```


Age
```{r}
fulldata$Age<-fulldata$Baseline.Age
```


Physical Activity Variables
* Note: Summed_PA_min is PER DAY 
* Walking, Moderate PA, and Vigorous PA values are MET min/week. IPAQ conversion is: Walking = 3.3 METs, Moderate PA = 4.0 METs, and Vigorous PA = 8.0 METs
```{r}
fulldata$Summed_PA_min<-fulldata$Summed.minutes.week.of.PA

fulldata$Summed_walking_min<-fulldata$MET.minutes.week.of.walking

fulldata$Summed_moderate_PA_min<-fulldata$MET.minutes.week.of.moderate.PA

fulldata$Summed_vigorous_PA_min<-fulldata$MET.minutes.week.of.vigorous.PA

fulldata$Summed_PA_MET_min<- (fulldata$Summed_walking_min + fulldata$Summed_moderate_PA_min + fulldata$Summed_vigorous_PA_min)
```


Index of Multiple Deprivation
```{r}
fulldata$Deprivation_Index<-NA
fulldata$Deprivation_Index[!is.na(fulldata$Index.of.multiple.deprivation.England) &
                          is.na(fulldata$Index.of.multiple.deprivation.Scotland) &   is.na(fulldata$Index.of.multiple.deprivation.Wales)]<-fulldata$Index.of.multiple.deprivation.England
fulldata$Deprivation_Index[is.na(fulldata$Index.of.multiple.deprivation.England) &
                          !is.na(fulldata$Index.of.multiple.deprivation.Scotland) &
is.na(fulldata$Index.of.multiple.deprivation.Wales)]<-fulldata$Index.of.multiple.deprivation.Scotland
fulldata$Deprivation_Index[is.na(fulldata$Index.of.multiple.deprivation.England) &
                          is.na(fulldata$Index.of.multiple.deprivation.Scotland) &
                          !is.na(fulldata$Index.of.multiple.deprivation.Wales)]<-fulldata$Index.of.multiple.deprivation.Wales
```


Neuropsychological testing variables
* "DSST" actually refers to Symbol Digit (not Digit Symbol) Substitution test
* For Pairs Matching, index 1 refers to the second round of the test, involving 6 pairs of cards.
```{r}
#Pairs Matching Score - this code corrected Jan. 9, 2025
fulldata$Pairs_matching_correct<-NA
fulldata$Pairs_matching_correct[!is.na(fulldata$Pairs.Matching.Correct.1)]<-fulldata$Pairs.Matching.Correct.1

fulldata$Pairs_matching_incorrect<-NA
fulldata$Pairs_matching_incorrect[!is.na(fulldata$Pairs.Matching.Incorrect.1)]<-fulldata$Pairs.Matching.Incorrect.1

#Trails B-A
fulldata$Trails_B_A<- fulldata$Trails.B.Time - fulldata$Trails.A.Time


#Symbol Digit Substitution Score
fulldata$Symbol.Digit<-fulldata$DSST.Total.Score

#Fluid Intelligence Score
fulldata$Fluid_intelligence<-fulldata$Fluid.Intellgence.Score

#Numeric Memory Score
fulldata$Numeric_memory<-fulldata$Numeric.Memory.Score
```


Greenspace and environmental variables (T0)
```{r}
#Greenspace percentage
fulldata$Greenspace_pcnt_1000m<-fulldata$Greenspace.percentage.1000m.0
fulldata$Greenspace_pcnt_300m<-fulldata$Greenspace.percentage.300m.0

#Natural Environment percentage
fulldata$Natural_environment_pcnt_1000m<-fulldata$Natural.environment.percentage.1000m.0
fulldata$Natural_environment_pcnt_300m<-fulldata$Natural.environment.percentage.300m.0
```

Brain volumes variables (2 timepoints [T2, T3])
* Using "Volumetric scaling from T1 image to standard space", we calculated HPC volumes adjusted for head size.
```{r}
#Grey matter volume (adjusted for head size)
fulldata$Grey_matter_volume_adjusted_T2<- fulldata$Grey.Matter.Volume..normalized.for.head.size..2
fulldata$Grey_matter_volume_adjusted_T3<- fulldata$Grey.Matter.Volume..normalized.for.head.size..3

#Grey matter volume (unadjusted)
fulldata$Grey_matter_volume_unadjusted_T2<- fulldata$Grey.Matter.Volume.unadjusted.2
fulldata$Grey_matter_volume_unadjusted_T3<- fulldata$Grey.Matter.Volume.unadjusted.3

#White matter volume (adjusted for head size)
fulldata$White_matter_volume_adjusted_T2<- fulldata$White.Matter.Volume..normalized.for.head.size..2
fulldata$White_matter_volume_adjusted_T3<- fulldata$White.Matter.Volume..normalized.for.head.size..3

#White matter volume (unadjusted)
fulldata$White_matter_volume_unadjusted_T2<- fulldata$White.Matter.Volume.unadjusted.2
fulldata$White_matter_volume_unadjusted_T3<- fulldata$White.Matter.Volume.unadjusted.3

#Total brain volume (adjusted for head size)
fulldata$Total_brain_volume_adjusted_T2 <- fulldata$Total.brain.volume..normalized.for.head.size..2
fulldata$Total_brain_volume_adjusted_T3 <- fulldata$Total.brain.volume..normalized.for.head.size..3

#Total brain volume (unadjusted)
fulldata$Total_brain_volume_adjusted_T2 <- fulldata$Total.brain.volume.unadjusted.2
fulldata$Total_brain_volume_adjusted_T3 <- fulldata$Total.brain.volume.unadjusted.3

#Hippocampal volume (Left Hemisphere; unadjusted)
fulldata$L_Hippocampal_volume_T2 <- fulldata$Left.hippocampal.volume.unadjusted.2
fulldata$L_Hippocampal_volume_T3 <- fulldata$Left.hippocampal.volume.unadjusted.2

#Hippocampal volume (Left Hemisphere; adjusted for head size)
fulldata$L_Hippocampal_volume_adjusted_T2 <- fulldata$L_Hippocampal_volume_T2 * fulldata$Volumetric.scaling.2
fulldata$L_Hippocampal_volume_adjusted_T3 <- fulldata$L_Hippocampal_volume_T3 * fulldata$Volumetric.scaling.3

#Hippocampal volume (Right Hemisphere; unadjusted)
fulldata$R_Hippocampal_volume_T2 <- fulldata$Right.hippocampal.volume.unadjusted.2
fulldata$R_Hippocampal_volume_T3 <- fulldata$Right.hippocampal.volume.unadjusted.3

#Hippocampal volume (Right Hemisphere; adjusted for head size)
fulldata$R_Hippocampal_volume_adjusted_T2 <- fulldata$R_Hippocampal_volume_T2 * fulldata$Volumetric.scaling.2
fulldata$R_Hippocampal_volume_adjusted_T3 <- fulldata$R_Hippocampal_volume_T3 * fulldata$Volumetric.scaling.3

#Average hippocampal volume (unadjusted)
fulldata$Avg_Hippocampal_volume_T2 <- (fulldata$L_Hippocampal_volume_T2 + fulldata$R_Hippocampal_volume_T2)/2
fulldata$Avg_Hippocampal_volume_T3 <- (fulldata$L_Hippocampal_volume_T3 + fulldata$R_Hippocampal_volume_T3)/2

#Average hippocampal volume (adjusted for head size)
fulldata$Avg_Hippocampal_volume_adjusted_T2 <- (fulldata$L_Hippocampal_volume_adjusted_T2 + fulldata$R_Hippocampal_volume_adjusted_T2)/2
fulldata$Avg_Hippocampal_volume_adjusted_T3 <- (fulldata$L_Hippocampal_volume_adjusted_T3 + fulldata$R_Hippocampal_volume_adjusted_T3)/2

#Total white matter hyperintensity volume
fulldata$Total_WMH_volume_T2<-fulldata$Total.volume.of.white.matter.hyperintensities.2
fulldata$Total_WMH_volume_T3<-fulldata$Total.volume.of.white.matter.hyperintensities.3
```


# 3) Determining final data set based on inclusion/exclusion criteria

First, only participants without stroke, Parkinson's, or dementia are included (N=483,664)
```{r}
included<-subset(fulldata, grepl("^\\s*$", Stroke.Date) & grepl("^\\s*$", Date.of.All.Cause.Dementia) & grepl("^\\s*$", Date.of.Parkinsons.Disease) &
                   CVD.problems!=3)
length(included$ID)
```

*Note: Counts for "included" and "excluded" do not sum to "fulldata" because missing (NA) values for CVD.problems will return false for both CVD.problems!=3 and CVD.problems==3. Therefore, adjust "included" to also include those with NA for CVD.problems.
```{r}
included2<-subset(fulldata, grepl("^\\s*$", Stroke.Date) & grepl("^\\s*$", Date.of.All.Cause.Dementia) & grepl("^\\s*$", Date.of.Parkinsons.Disease) & (CVD.problems!=3 | is.na(CVD.problems)))
length(included2$ID)
```

Excluded participants (N=17,863)
```{r}
Excluded<-subset(fulldata, !grepl("^\\s*$", Stroke.Date) | !grepl("^\\s*$", Date.of.All.Cause.Dementia) | !grepl("^\\s*$", Date.of.Parkinsons.Disease) |
                   CVD.problems==3)
length(Excluded$ID)
```


For the combined cognitive and brain analysis, only participants with complete cognitive and brain data were included (N= 17,494)
```{r}
CogsBrain<- subset(included, !is.na(Grey_matter_volume_adjusted_T2) & !is.na(White_matter_volume_adjusted_T2) &
 !is.na(Avg_Hippocampal_volume_adjusted_T2) &
                  !is.na(Trails_B_A) & !is.na(Symbol.Digit) &
                 !is.na(Greenspace.percentage.1000m.0) & !is.na(Greenspace.percentage.300m.0) &
                  !is.na(Natural.environment.percentage.1000m.0) & !is.na(Natural.environment.percentage.300m.0))
length(CogsBrain$ID)
```


For the final analytical sample, only participants with complete cognitive, brain, and covariate data were included (N=11,448)
* Observations were excluded if any values for the above were: NA, "Do not know or prefer not to answer","Do not know or prefer not to say","Don't know or prefer not to answer","Prefer not to answer", "Do not know", "Postcode not linkable".
```{r}
#include observations if the values are not missing
CogsBrainCov <- included[!apply(included[c("Grey_matter_volume_adjusted_T2","White_matter_volume_adjusted_T2","Trails_B_A","Symbol.Digit","Greenspace.percentage.1000m.0","Greenspace.percentage.300m.0","Natural.environment.percentage.1000m.0","Natural.environment.percentage.300m.0","Age","Sex","BMI","Education","Income","Average_time_outdoors","Summed_PA_min","Smoking","Disability","Deprivation_Index","Population_density")], 1, function(row) any(row %in% c("Do not know or prefer not to answer","Do not know or prefer not to say","Don't know or prefer not to answer","Prefer not to answer", "Do not know", "Postcode not linkable") | is.na(row))), ]
length(CogsBrainCov$ID)
```

*Adjusted final analytical sample, starting from "included2". Luckily, this leads to the same final sample as that derived from "included", so no further adjustments need to be made.
```{r}
#include observations if the values are not missing
CogsBrainCov2 <- included2[!apply(included2[c("Grey_matter_volume_adjusted_T2","White_matter_volume_adjusted_T2","Trails_B_A","Symbol.Digit","Greenspace.percentage.1000m.0","Greenspace.percentage.300m.0","Natural.environment.percentage.1000m.0","Natural.environment.percentage.300m.0","Age","Sex","BMI","Education","Income","Average_time_outdoors","Summed_PA_min","Smoking","Disability","Deprivation_Index","Population_density")], 1, function(row) any(row %in% c("Do not know or prefer not to answer","Do not know or prefer not to say","Don't know or prefer not to answer","Prefer not to answer", "Do not know", "Postcode not linkable") | is.na(row))), ]
length(CogsBrainCov2$ID)
```

Check distribution of missing data across nature exposure, brain structure, cognitive functions, and covariates
```{r}
#missing nature exposure
missingNature <- included2[apply(included2[c("Greenspace.percentage.1000m.0","Greenspace.percentage.300m.0","Natural.environment.percentage.1000m.0","Natural.environment.percentage.300m.0")], 1, function(row) any(row %in% c("Do not know or prefer not to answer","Do not know or prefer not to say","Don't know or prefer not to answer","Prefer not to answer", "Do not know", "Postcode not linkable") | is.na(row))), ]
length(missingNature$ID)

#missing brain structure
missingBrain <- included2[apply(included2[c("Grey_matter_volume_adjusted_T2","White_matter_volume_adjusted_T2")], 1, function(row) any(row %in% c("Do not know or prefer not to answer","Do not know or prefer not to say","Don't know or prefer not to answer","Prefer not to answer", "Do not know", "Postcode not linkable") | is.na(row))), ]
length(missingBrain$ID)

#missing cognitive function
missingCog <- included2[apply(included2[c("Trails_B_A","Symbol.Digit")], 1, function(row) any(row %in% c("Do not know or prefer not to answer","Do not know or prefer not to say","Don't know or prefer not to answer","Prefer not to answer", "Do not know", "Postcode not linkable") | is.na(row))), ]
length(missingCog$ID)

#missing covariates
missingCovars <- included2[apply(included2[c("Age","Sex","BMI","Education","Income","Average_time_outdoors","Summed_PA_min","Smoking","Disability","Deprivation_Index","Population_density")], 1, function(row) any(row %in% c("Do not know or prefer not to answer","Do not know or prefer not to say","Don't know or prefer not to answer","Prefer not to answer", "Do not know", "Postcode not linkable") | is.na(row))), ]
length(missingCovars$ID)
```


These are the simplified data sets we'll be using for 1) the full data set (i.e., all ~500,000 participants), 2) participants with both cognitive and brain data (N= 17,494), and 3) participants with cognitive, brain, and covariate data (N= 11,448).
```{r}
Full.df<-fulldata[c(1,113,4,11,112,94,5,6,108,109,106,107,102,103,101,105,100,111,104,114,115,118,97:99,119,110,122,123,126:129,130,131,134,135,138,139,150:153,91:93,41,43,44)]

CogsBrain.df<-CogsBrain[c(1,113,4,11,112,94,5,6,108,109,106,107,102,103,101,105,100,111,104,114,115,118,97:99,119,110,122,123,126:129,130,131,134,135,138,139,150:153,91:93,41,43,44)]

CogsBrainCov.df<-CogsBrainCov[c(1,113,4,11,112,94,5,6,108,109,106,107,102,103,101,105,100,111,104,114,115,118,97:99,119,110,122,123,126:129,130,131,134,135,138,139,150:153,91:93,41,43,44)]

#Adding the other 3 cognitive function tests
CogsPlusBrainCov.df <- CogsBrainCov[c(1,113,4,11,112,94,5,6,108,109,106,107,102,103,101,105,100,111,104,114,115,118,97:99,119,110,122,123,124,125,121,126:129,130,131,134,135,138,139,150:153,91:93,41,43,44)]
```


This step creates variable labels that are easy to understand in the data output.
```{r}
Full.df = apply_labels(Full.df,
                       Age = "Age (years)",
                       BMI = "Body Mass Index (kg/m^2)",
                       Income = "Household Income",
                       Time_outdoors_summer_cont = "Time outdoors in the summer (hours/day)",
                       Time_outdoors_winter_cont = "Time outdoors in the winter (hours/day)",
                       Average_time_outdoors = "Average time outdoors (hours/day)",
                       Hearing_difficulty = "Hearing problems",
                       Eye_problems = "Eye problems",
                       Disability = "Living with a disability or long-term illness",
                       Diabetes = "Living with diabetes mellitus",
                       Cancer = "Living with or previously had cancer",
                       Other_medical_conditions = "Living with another serious medical condition",
                       CVD_problems = "Living with cardiovascular disease",
                       Clots_Asthma_Allergies = "Living with blood clots, allergies, or asthma",
                       Bipolar_Depression = "Living with Bipolar Disorder or Depression",
                       Alcohol = "Alcohol drinking status",
                       Smoking = "Smoking status",
                       Population_density = "Home area population density",
                       Summed_PA_min = "Total physical activity (min/day)",
                       Summed_walking_min = "Total walking (MET min/week)",
                       Summed_PA_MET_min = "Total physical activity (MET min/week)",
                       Trails_B_A = "Trail making test (B-A; sec)",
                       Symbol.Digit = "Symbol digit substitution test (correct matches)",
                       Greenspace_pcnt_300m = "Greenspace percentage (buffered to 300 meters)",
                       Greenspace_pcnt_1000m = "Greenspace percentage (buffered to 1000 meters)",
                       Natural_environment_pcnt_300m = "Natural environment percentage (buffered to 300 meters)",
                       Natural_environment_pcnt_1000m = "Natural environment percentage (buffered to 1000 meters)",
                       Deprivation_Index = "Townsend Deprivation Index",
                       Grey_matter_volume_adjusted_T2 = "Grey matter volume at imaging study entry (adjusted for head size; mm^3)",
                       Grey_matter_volume_adjusted_T3  = "Grey matter volume at imaging study follow-up (adjusted for head size; mm^3)",
                       White_matter_volume_adjusted_T2 = "White matter volume at imaging study entry (adjusted for head size; mm^3)",
                       White_matter_volume_adjusted_T3 = "White matter volume at imaging study follow-up (adjusted for head size; mm^3)",
                       Avg_Hippocampal_volume_adjusted_T2 = "Average hippocampal volume at imaging study entry (adjusted for head size; mm^3)",
                       Avg_Hippocampal_volume_adjusted_T3 ="Average hippocampal volume at imaging study follow-up (adjusted for head size; mm^3)",
                       Total_WMH_volume_T2 = "White matter hyperintensities volume at imaging study entry (mm^3)",
                       Total_WMH_volume_T3 = "White matter hyperintensities volume at imaging study follow-up (mm^3)"
)

CogsBrainCov.df = apply_labels(CogsBrainCov.df,
                       Age = "Age (years)",
                       BMI = "Body Mass Index (kg/m^2)",
                       Income = "Household Income",
                       Time_outdoors_summer_cont = "Time outdoors in the summer (hours/day)",
                       Time_outdoors_winter_cont = "Time outdoors in the winter (hours/day)",
                       Average_time_outdoors = "Average time outdoors (hours/day)",
                       Hearing_difficulty = "Hearing problems",
                       Eye_problems = "Eye problems",
                       Disability = "Living with a disability or long-term illness",
                       Diabetes = "Living with diabetes mellitus",
                       Cancer = "Living with or previously had cancer",
                       Other_medical_conditions = "Living with another serious medical condition",
                       CVD_problems = "Living with cardiovascular disease",
                       Clots_Asthma_Allergies = "Living with blood clots, allergies, or asthma",
                       Bipolar_Depression = "Living with Bipolar Disorder or Depression",
                       Alcohol = "Alcohol drinking status",
                       Smoking = "Smoking status",
                       Population_density = "Home area population density",
                       Summed_PA_min = "Total physical activity (min/day)",
                       Summed_walking_min = "Total walking (MET min/week)",
                       Summed_PA_MET_min = "Total physical activity (MET min/week)",
                       Trails_B_A = "Trail making test (B-A; sec)",
                       Symbol.Digit = "Symbol digit substitution test (correct matches)",
                       Greenspace_pcnt_300m = "Greenspace percentage (buffered to 300 meters)",
                       Greenspace_pcnt_1000m = "Greenspace percentage (buffered to 1000 meters)",
                       Natural_environment_pcnt_300m = "Natural environment percentage (buffered to 300 meters)",
                       Natural_environment_pcnt_1000m = "Natural environment percentage (buffered to 1000 meters)",
                       Deprivation_Index = "Townsend Deprivation Index",
                       Grey_matter_volume_adjusted_T2 = "Grey matter volume at imaging study entry (adjusted for head size; mm^3)",
                       Grey_matter_volume_adjusted_T3  = "Grey matter volume at imaging study follow-up (adjusted for head size; mm^3)",
                       White_matter_volume_adjusted_T2 = "White matter volume at imaging study entry (adjusted for head size; mm^3)",
                       White_matter_volume_adjusted_T3 = "White matter volume at imaging study follow-up (adjusted for head size; mm^3)",
                       Avg_Hippocampal_volume_adjusted_T2 = "Average hippocampal volume at imaging study entry (adjusted for head size; mm^3)",
                       Avg_Hippocampal_volume_adjusted_T3 ="Average hippocampal volume at imaging study follow-up (adjusted for head size; mm^3)",
                       Total_WMH_volume_T2 = "White matter hyperintensities volume at imaging study entry (mm^3)",
                       Total_WMH_volume_T3 = "White matter hyperintensities volume at imaging study follow-up (mm^3)"
)

CogsPlusBrainCov.df = apply_labels(CogsPlusBrainCov.df,
                       Age = "Age (years)",
                       BMI = "Body Mass Index (kg/m^2)",
                       Income = "Household Income",
                       Time_outdoors_summer_cont = "Time outdoors in the summer (hours/day)",
                       Time_outdoors_winter_cont = "Time outdoors in the winter (hours/day)",
                       Average_time_outdoors = "Average time outdoors (hours/day)",
                       Hearing_difficulty = "Hearing problems",
                       Eye_problems = "Eye problems",
                       Disability = "Living with a disability or long-term illness",
                       Diabetes = "Living with diabetes mellitus",
                       Cancer = "Living with or previously had cancer",
                       Other_medical_conditions = "Living with another serious medical condition",
                       CVD_problems = "Living with cardiovascular disease",
                       Clots_Asthma_Allergies = "Living with blood clots, allergies, or asthma",
                       Bipolar_Depression = "Living with Bipolar Disorder or Depression",
                       Alcohol = "Alcohol drinking status",
                       Smoking = "Smoking status",
                       Population_density = "Home area population density",
                       Summed_PA_min = "Total physical activity (min/day)",
                       Summed_walking_min = "Total walking (MET min/week)",
                       Summed_PA_MET_min = "Total physical activity (MET min/week)",
                       Trails_B_A = "Trail making test (B-A; sec)",
                       Symbol.Digit = "Symbol digit substitution test (correct matches)",
                       Fluid_intelligence = "Fluid intelligence (correct answers)",
                       Numeric_memory = "Numeric memory (maximum digits recalled)",
                       Pairs_matching_incorrect = "Pairs matching (incorrect matches)",
                       Greenspace_pcnt_300m = "Greenspace percentage (buffered to 300 meters)",
                       Greenspace_pcnt_1000m = "Greenspace percentage (buffered to 1000 meters)",
                       Natural_environment_pcnt_300m = "Natural environment percentage (buffered to 300 meters)",
                       Natural_environment_pcnt_1000m = "Natural environment percentage (buffered to 1000 meters)",
                       Deprivation_Index = "Townsend Deprivation Index",
                       Grey_matter_volume_adjusted_T2 = "Grey matter volume at imaging study entry (adjusted for head size; mm^3)",
                       Grey_matter_volume_adjusted_T3  = "Grey matter volume at imaging study follow-up (adjusted for head size; mm^3)",
                       White_matter_volume_adjusted_T2 = "White matter volume at imaging study entry (adjusted for head size; mm^3)",
                       White_matter_volume_adjusted_T3 = "White matter volume at imaging study follow-up (adjusted for head size; mm^3)",
                       Avg_Hippocampal_volume_adjusted_T2 = "Average hippocampal volume at imaging study entry (adjusted for head size; mm^3)",
                       Avg_Hippocampal_volume_adjusted_T3 ="Average hippocampal volume at imaging study follow-up (adjusted for head size; mm^3)",
                       Total_WMH_volume_T2 = "White matter hyperintensities volume at imaging study entry (mm^3)",
                       Total_WMH_volume_T3 = "White matter hyperintensities volume at imaging study follow-up (mm^3)"
)
```


# 4) Compare the complete data set with the final cognitive+brain+covariate data set

This step will summarize participant characteristics for the full data set, as compared to the final analytical data set
```{r}
Full.df$included<-"All participants"
CogsBrain.df$included<-"Participants with both brain and cognitive data" 
CogsBrainCov.df$included<-"Participants with brain, cognitive, and covariate data"

Tableone.df<-rbind(Full.df,CogsBrainCov.df)

vars<- Tableone.df %>% select(c(2:43,50))

vars %>% tbl_summary(
  statistic = list(all_continuous() ~ "{mean} ({sd})",
                   all_categorical() ~ "{n} ({p}%)"),
  digits = all_continuous() ~ 2,
  by = included
)
```

Table for cognitive+brain+covariate subset, stratified by sex
```{r}
tbl_summary(
  select(CogsBrainCov.df, c(2:43)),
  by = "Sex", 
  statistic = list(all_continuous() ~ "{mean} ({sd})",
                   all_categorical() ~ "{n} ({p}%)"),
  digits = all_continuous() ~ 2
) %>%
  add_p(test = list(all_continuous() ~ "t.test",
                   all_categorical() ~ "chisq.test"))
```

# 5) Regression analyses examining the associations of natural environment with cognitive function [subset having cognitive, brain, and covariate data]

First, this step recodes our covariates of interest into scales which are easier to interpret in a regression. Important notes:
1. Every unit of age represents 10 years above the mean age
2. Every unit of BMI represents 5 kg/m^2 above the mean BMI
3. Every unit of natural environment represents a %10 increase in natural environment
4. Brain volumes are reformatted to cubic centimeters (cm^3)
5. Education kept as nominal.
```{r}
library(dplyr)

CogsBrainCov.df2<-CogsBrainCov.df %>%
  mutate(Age = scale(Age, scale=FALSE)/10,
         Sex = recode(Sex,
                      "Female" = 0.5,
                      "Male" = -0.5),
         BMI = scale(BMI, scale=FALSE)/5,
         Race = recode(Race,
                       "White" = 0.5,
                       "Mixed" = -0.5,
                       "Black or Black British"= -0.5,
                       "Asian or Asian British"= -0.5),
         Income = as.numeric(factor(Income,
                                    levels = c("<18,000 GBP","18,000-30,999 GBP","31,000-51,999 GBP","52,000-100,000 GBP",">100,000 GBP"))),
         Disability = recode(Disability, 
                             "Yes" = 1,
                             "No" = 0),
         Diabetes = recode(Diabetes,
                           "Yes" = 1,
                           "No" = 0),
         CVD_problems = recode(CVD_problems,
                               "Yes" = 1,
                               "No" = 0),
         Cancer = recode(Cancer,
                         "Yes" = 1,
                         "No" = 0),
         Other_medical_conditions =  recode(Other_medical_conditions,
                                            "Yes" = 1,
                                            "No" = 0),
         Alcohol = as.numeric(factor(Alcohol,
                                     levels = c("Never","Previous drinker","Current drinker"))),
         Smoking = as.numeric(factor(Smoking,
                                     levels = c("Never", "Previous Smoker", "Current Smoker"))),
         Population_density = as.numeric(factor(Population_density,
                                                levels = c("Urban area", "Town", "Village, Hamlet, or Rural Area"))),
         Natural_environment_1000m_10pcnt = Natural_environment_pcnt_1000m/10,
         Greenspace_1000m_10pcnt = Greenspace_pcnt_1000m/10,
         Natural_environment_300m_10pcnt = Natural_environment_pcnt_300m/10,
         Greenspace_300m_10pcnt = Greenspace_pcnt_300m/10,
         Grey_matter_volume_adjusted_T2_cm3 = Grey_matter_volume_adjusted_T2/1000,
         White_matter_volume_adjusted_T2_cm3 = White_matter_volume_adjusted_T2/1000,
         Avg_Hippocampal_volume_adjusted_T2_cm3 = Avg_Hippocampal_volume_adjusted_T2/1000,
         Total_WMH_volume_T2_cm3 = Total_WMH_volume_T2/1000
         )

detach(package:dplyr,unload=TRUE)
```

Repeating above step for separate dataset with the additional cognitive outcomes
```{r}
library(dplyr)

CogsPlusBrainCov.df2<-CogsPlusBrainCov.df %>%
  mutate(Age = scale(Age, scale=FALSE)/10,
         Sex = recode(Sex,
                      "Female" = 0.5,
                      "Male" = -0.5),
         BMI = scale(BMI, scale=FALSE)/5,
         Race = recode(Race,
                       "White" = 0.5,
                       "Mixed" = -0.5,
                       "Black or Black British"= -0.5,
                       "Asian or Asian British"= -0.5),
         Income = as.numeric(factor(Income,
                                    levels = c("<18,000 GBP","18,000-30,999 GBP","31,000-51,999 GBP","52,000-100,000 GBP",">100,000 GBP"))),
         Disability = recode(Disability, 
                             "Yes" = 1,
                             "No" = 0),
         Diabetes = recode(Diabetes,
                           "Yes" = 1,
                           "No" = 0),
         CVD_problems = recode(CVD_problems,
                               "Yes" = 1,
                               "No" = 0),
         Cancer = recode(Cancer,
                         "Yes" = 1,
                         "No" = 0),
         Other_medical_conditions =  recode(Other_medical_conditions,
                                            "Yes" = 1,
                                            "No" = 0),
         Alcohol = as.numeric(factor(Alcohol,
                                     levels = c("Never","Previous drinker","Current drinker"))),
         Smoking = as.numeric(factor(Smoking,
                                     levels = c("Never", "Previous Smoker", "Current Smoker"))),
         Population_density = as.numeric(factor(Population_density,
                                                levels = c("Urban area", "Town", "Village, Hamlet, or Rural Area"))),
         Natural_environment_1000m_10pcnt = Natural_environment_pcnt_1000m/10,
         Greenspace_1000m_10pcnt = Greenspace_pcnt_1000m/10,
         Natural_environment_300m_10pcnt = Natural_environment_pcnt_300m/10,
         Greenspace_300m_10pcnt = Greenspace_pcnt_300m/10,
         Grey_matter_volume_adjusted_T2_cm3 = Grey_matter_volume_adjusted_T2/1000,
         White_matter_volume_adjusted_T2_cm3 = White_matter_volume_adjusted_T2/1000,
         Avg_Hippocampal_volume_adjusted_T2_cm3 = Avg_Hippocampal_volume_adjusted_T2/1000,
         Total_WMH_volume_T2_cm3 = Total_WMH_volume_T2/1000
         )

detach(package:dplyr,unload=TRUE)
```

Define function to get both model summary and confidence interval
```{r}
summary_confint <- function(model) {
  list(summary(model), confint(model))
  }
```

Define function for regression model assumption checks
```{r}
# Install package for Durbin-Watson test
options(repos = c(CRAN = "https://cran.r-project.org"))
install.packages(c("lmtest", "gvlma"))
library("lmtest")
library("gvlma")

model_diagnostics <- function(model, bin_width=1) {
  # Residuals vs. Fitted plot for linearity, normality, equivariance
  # Normal Q-Q plot for normality
  # Residuals vs. Leverage plot for influential outliers
  plot(model, which = c(1,2,5))

  # Histogram of residuals for normality
  hist(residuals(model), breaks = seq(floor(min(residuals(model))), ceiling(max(residuals(model))), by = bin_width))
  
  # Durbin-Watson test for independent errors
  print(dwtest(model))
  
  # Test skewness, kurtosis, equivariance
  gvlma(model)
  
  # Test multicollinearity among predictors (in interaction models) and among covariates?
  #vif(model)
}
```


## 5.1) 1000m buffer for natural environment percentage

### 5.1.1) Models for the association of natural environment with trail making test (B-A)

Crude model (Covariates: Age, Sex, BMI, Education, and Income)
```{r}
summary(lm(Trails_B_A~Natural_environment_1000m_10pcnt + Age + Sex + BMI + Education + Income, CogsBrainCov.df2))
```

Full adjusted model (Covariates: Age, Sex, BMI, Education, Income, Average time spent outdoors, Physical activity minutes, Smoking status, Disability status, Townsend Deprivation Index, and Population Density)
```{r}
model5b1 <- lm(Trails_B_A~Natural_environment_1000m_10pcnt + Age + Sex + BMI + Education + Income + Average_time_outdoors +  
             Summed_PA_min + Smoking + Disability + Deprivation_Index + Population_density, CogsBrainCov.df2)
summary(model5b1)
confint(model5b1,"Natural_environment_1000m_10pcnt") #95% confidence interval
```


Model assumption checks
```{r}
model_diagnostics(model5b1)

  # Residuals vs. Fitted plot suggests approximately linear, minor heteroskedasticity
  # Normal Q-Q plot suggests non-normal skew
  # Residuals vs. Leverage plot: no outliers of major concern
  # Histogram of residuals: approximately normal, maybe slight positive skew
  # Durbin-Watson test near 2 suggests independent residuals
  # Skewness, Kurtosis, and Heteroskedasticity tests suggest non-normal and heteroskedastic
```

### 5.1.2) Models for the association of natural environment with Symbol Digit Substitution test

Crude model (Covariates: Age, Sex, BMI, Education, and Income)
```{r}
summary(lm(Symbol.Digit~Natural_environment_1000m_10pcnt + Age + Sex + BMI + Education + Income, CogsBrainCov.df2))
```

Full adjusted model (Covariates: Age, Sex, BMI, Education, Income, Average time spent outdoors, Physical activity minutes, Smoking status, Disability status, Townsend Deprivation Index, and Population Density)
```{r}
model5b2 <- lm(Symbol.Digit~Natural_environment_1000m_10pcnt + Age + Sex + BMI + Education + Income + Average_time_outdoors + 
             Summed_PA_min + Smoking + Disability + Deprivation_Index + Population_density, CogsBrainCov.df2)
summary(model5b2)
confint(model5b2,"Natural_environment_1000m_10pcnt") #95% confidence interval
```

Line graph illustration of the full adjusted model
* Note: change of natural environment variable (not divided into 10% increments)
```{r}
mdl.3<-resid(lm(Symbol.Digit~Age + Sex + BMI + Education + Income + 
                  Average_time_outdoors + 
                  Summed_PA_min + Smoking + Disability + Deprivation_Index + Population_density, CogsBrainCov.df2))

mdl.4<-resid(lm(Natural_environment_pcnt_1000m~Age + Sex + BMI + Education + Income + 
                  Average_time_outdoors + 
                  Summed_PA_min + Smoking + Disability + Deprivation_Index + Population_density, CogsBrainCov.df2))

Symbol.Digit.graph<-as.data.frame(cbind(mdl.3,mdl.4))
Symbol.Digit.graph$mdl.3<-Symbol.Digit.graph$mdl.3 + mean(CogsBrainCov.df2$Symbol.Digit)
Symbol.Digit.graph$mdl.4<-Symbol.Digit.graph$mdl.4 + mean(CogsBrainCov.df2$Natural_environment_pcnt_1000m)


ggplot(data = Symbol.Digit.graph, aes(x = mdl.4, y = mdl.3)) + 
  labs(x="Natural Environment (%)", y="Symbol Digit Substitution Test (correct matches)")  + geom_smooth(method= 'lm',  se = TRUE, color= 'black') +
  xlim(0,100) +
  theme_bw() + theme(panel.border = element_blank(), panel.grid.major = element_blank(),
                     panel.grid.minor = element_blank(), axis.line = element_line(colour = "black"))
```

Model assumption checks
```{r}
model_diagnostics(model5b2)

  # Residuals vs. Fitted plot suggests approximately linear, some heteroskedasticity, 1 outlier
  # Normal Q-Q plot suggests non-normal skew, 1 outlier
  # Residuals vs. Leverage plot: 1 clear outlier
  # Histogram of residuals: approximately normal, slight negative skew
  # Durbin-Watson test near 2 suggests independent residuals
  # Skewness, Kurtosis, and Heteroskedasticity tests suggest non-normal but homoskedastic
```


## 5.2) 300m buffer for natural environment percentage

### 5.2.1) Association of natural environment with Trail-making Test (B-A) (not significant)

Crude model (Covariates: Age, Sex, BMI, Education, and Income)
```{r}
summary(lm(Trails_B_A~Natural_environment_300m_10pcnt + Age + Sex + BMI + Education + Income, CogsBrainCov.df2))
```

Full adjusted model (Covariates: Age, Sex, BMI, Education, Income, Average time spent outdoors, Physical activity minutes, Smoking status, Disability status, Townsend Deprivation Index, and Population Density)
```{r}
summary_confint(lm(Trails_B_A~Natural_environment_300m_10pcnt + Age + Sex + BMI + Education + Income + Average_time_outdoors + Summed_PA_min + Smoking + Disability + Deprivation_Index + Population_density, CogsBrainCov.df2))
```

Model assumption checks
```{r}
model_diagnostics(lm(Trails_B_A~Natural_environment_300m_10pcnt + Age + Sex + BMI + Education + Income + Average_time_outdoors + Summed_PA_min + Smoking + Disability + Deprivation_Index + Population_density, CogsBrainCov.df2))

  # Residuals vs. Fitted plot 
  # Normal Q-Q plot 
  # Residuals vs. Leverage plot
  # Histogram of residuals
  # Durbin-Watson test 
  # Skewness, Kurtosis, and Heteroskedasticity tests 
```

### 5.2.2) Association of natural environment with Symbol Digit Substitution (significant)

Crude model (Covariates: Age, Sex, BMI, Education, and Income)
```{r}
summary_confint(lm(Symbol.Digit~Natural_environment_300m_10pcnt + Age + Sex + BMI + Education + Income, CogsBrainCov.df2))
```

Full adjusted model (Covariates: Age, Sex, BMI, Education, Income, Average time spent outdoors, Physical activity minutes, Smoking status, Disability status, Townsend Deprivation Index, and Population Density)
```{r}
summary_confint(lm(Symbol.Digit~Natural_environment_300m_10pcnt + Age + Sex + BMI + Education + Income + 
             Average_time_outdoors + 
             Summed_PA_min + Smoking + Disability + Deprivation_Index + Population_density, CogsBrainCov.df2))
```

Model assumption checks
```{r}
model_diagnostics(lm(Symbol.Digit~Natural_environment_300m_10pcnt + Age + Sex + BMI + Education + Income + Average_time_outdoors + Summed_PA_min + Smoking + Disability + Deprivation_Index + Population_density, CogsBrainCov.df2))

  # Residuals vs. Fitted plot 
  # Normal Q-Q plot 
  # Residuals vs. Leverage plot
  # Histogram of residuals
  # Durbin-Watson test 
  # Skewness, Kurtosis, and Heteroskedasticity tests 
```

# 6) Regression analyses examining the association of natural environment with brain volumes [cognitive+brain+covariate subset]

## 6.1) 1000m natural environment buffer

### 6.1.1) Models for the association of natural environment with grey matter volume

Crude model (Covariates: Age, Sex, BMI, Education, and Income)
```{r}
summary(lm(Grey_matter_volume_adjusted_T2_cm3~Natural_environment_1000m_10pcnt + Age + Sex + BMI + Education + Income, CogsBrainCov.df2))
```

Full adjusted model (Covariates: Age, Sex, BMI, Education, Income, Average time spent outdoors, Physical activity minutes, Smoking status, Disability status, Townsend Deprivation Index, and Population Density)
```{r}
model6b1 <- lm(Grey_matter_volume_adjusted_T2_cm3~Natural_environment_1000m_10pcnt + Age + Sex + BMI + Education + Income + Average_time_outdoors + Summed_PA_min + Smoking + Disability + Deprivation_Index + Population_density, CogsBrainCov.df2)
summary(model6b1)
confint(model6b1,"Natural_environment_1000m_10pcnt") #95% confidence interval
```

Line graph illustration of the full adjusted model
```{r}
mdl.5b<-resid(lm(Grey_matter_volume_adjusted_T2_cm3~Age + Sex + BMI + Education + Income + 
                  Average_time_outdoors + 
                  Summed_PA_min + Smoking + Disability + Deprivation_Index + Population_density, CogsBrainCov.df2))

mdl.6b<-resid(lm(Natural_environment_1000m_10pcnt~Age + Sex + BMI + Education + Income + 
                  Average_time_outdoors + 
                  Summed_PA_min + Smoking + Disability + Deprivation_Index + Population_density, CogsBrainCov.df2))

GMV.graph<-as.data.frame(cbind(mdl.5b,mdl.6b))
GMV.graph$mdl.5b<-GMV.graph$mdl.5b + mean(CogsBrainCov.df2$Grey_matter_volume_adjusted_T2_cm3)
GMV.graph$mdl.6b<-GMV.graph$mdl.6b*10 + mean(CogsBrainCov.df2$Natural_environment_1000m_10pcnt)*10


ggplot(data = GMV.graph, aes(x = mdl.6b, y = mdl.5b)) + 
  labs(x="Natural Environment (%)", y="Grey Matter Volume (cm^3)")  + geom_smooth(method= 'lm',  se = TRUE, color= 'black') +
  coord_cartesian(xlim=c(0,100)) +
  theme_bw() + theme(panel.border = element_blank(), panel.grid.major = element_blank(),
                     panel.grid.minor = element_blank(), axis.line = element_line(colour = "black"))
```

Model assumption checks
```{r}
model_diagnostics(model6b1)
```

### 6.1.2) Models for the association of natural environment with white matter volume

Crude model (Covariates: Age, Sex, BMI, Education, and Income)
```{r}
summary(lm(White_matter_volume_adjusted_T2_cm3~Natural_environment_1000m_10pcnt + Age + Sex + BMI + Education + Income, CogsBrainCov.df2))
```

Full adjusted model (Covariates: Age, Sex, BMI, Education, Income, Average time spent outdoors, Physical activity minutes, Smoking status, Disability status, Townsend Deprivation Index, and Population Density)
```{r}
model6b2 <- lm(White_matter_volume_adjusted_T2_cm3~Natural_environment_1000m_10pcnt + Age + Sex + BMI + Education + Income + 
             Average_time_outdoors + 
             Summed_PA_min + Smoking + Disability + Deprivation_Index + Population_density, CogsBrainCov.df2)
summary(model6b2)
confint(model6b2,"Natural_environment_1000m_10pcnt") #95% confidence interval
```

Line graph illustration of the full adjusted model
```{r}
mdl.7b<-resid(lm(White_matter_volume_adjusted_T2_cm3~Age + Sex + BMI + Education + Income + 
                  Average_time_outdoors + 
                  Summed_PA_min + Smoking + Disability + Deprivation_Index + Population_density, CogsBrainCov.df2))

mdl.8b<-resid(lm(Natural_environment_1000m_10pcnt~Age + Sex + BMI + Education + Income + 
                  Average_time_outdoors + 
                  Summed_PA_min + Smoking + Disability + Deprivation_Index + Population_density, CogsBrainCov.df2))

WMV.graph<-as.data.frame(cbind(mdl.7b,mdl.8b))
WMV.graph$mdl.7b<-WMV.graph$mdl.7b + mean(CogsBrainCov.df2$White_matter_volume_adjusted_T2_cm3)
WMV.graph$mdl.8b<-WMV.graph$mdl.8b*10 + mean(CogsBrainCov.df2$Natural_environment_1000m_10pcnt)*10


ggplot(data = WMV.graph, aes(x = mdl.8b, y = mdl.7b)) + 
  labs(x="Natural Environment (%)", y="White Matter Volume (cm^3)")  + geom_smooth(method= 'lm',  se = TRUE, color= 'black') +
  xlim(0,100) +
  theme_bw() + theme(panel.border = element_blank(), panel.grid.major = element_blank(),
                     panel.grid.minor = element_blank(), axis.line = element_line(colour = "black"))

```

Model assumption checks
```{r}
model_diagnostics(model6b2)
```

### 6.1.3) Models for the association of natural environment with hippocampal volume

Crude model (Covariates: Age, Sex, BMI, Education, and Income)
```{r}
summary(lm(Avg_Hippocampal_volume_adjusted_T2_cm3~Natural_environment_1000m_10pcnt + Age + Sex + BMI + Education + Income, CogsBrainCov.df2))
```

Full adjusted model (Covariates: Age, Sex, BMI, Education, Income, Average time spent outdoors, Physical activity minutes, Smoking status, Disability status, Townsend Deprivation Index, and Population Density)
```{r}
model6b3 <- lm(Avg_Hippocampal_volume_adjusted_T2_cm3~Natural_environment_1000m_10pcnt + Age + Sex + BMI + Education + Income + 
             Average_time_outdoors + 
             Summed_PA_min + Smoking + Disability + Deprivation_Index + Population_density, CogsBrainCov.df2)
summary(model6b3)
confint(model6b3,"Natural_environment_1000m_10pcnt") #95% confidence interval
```

Model assumption checks
```{r}
model_diagnostics(model6b3)
```

## 6.2) 300m natural environment buffer

### 6.2.1) Association of natural environment with grey matter volume (significant)

Crude model (Covariates: Age, Sex, BMI, Education, and Income)
```{r}
summary_confint(lm(Grey_matter_volume_adjusted_T2_cm3~Natural_environment_300m_10pcnt + Age + Sex + BMI + Education + Income, CogsBrainCov.df2))
```

Full adjusted model (Covariates: Age, Sex, BMI, Education, Income, Average time spent outdoors, Physical activity minutes, Smoking status, Disability status, Townsend Deprivation Index, and Population Density)
```{r}
summary_confint(lm(Grey_matter_volume_adjusted_T2_cm3~Natural_environment_300m_10pcnt + Age + Sex + BMI + Education + Income + Average_time_outdoors + Summed_PA_min + Smoking + Disability + Deprivation_Index + Population_density, CogsBrainCov.df2))
```

### 6.2.2) Association of natural environment with white matter volume (significant)

Crude model (Covariates: Age, Sex, BMI, Education, and Income)
```{r}
summary_confint(lm(White_matter_volume_adjusted_T2_cm3~Natural_environment_300m_10pcnt + Age + Sex + BMI + Education + Income, CogsBrainCov.df2))
```

Full adjusted model (Covariates: Age, Sex, BMI, Education, Income, Average time spent outdoors, Physical activity minutes, Smoking status, Disability status, Townsend Deprivation Index, and Population Density)
```{r}
summary_confint(lm(White_matter_volume_adjusted_T2_cm3~Natural_environment_300m_10pcnt + Age + Sex + BMI + Education + Income + Average_time_outdoors + Summed_PA_min + Smoking + Disability + Deprivation_Index + Population_density, CogsBrainCov.df2))
```

### 6.2.3) Association of natural environment with hippocampal volume (not significant)

Crude model (Covariates: Age, Sex, BMI, Education, and Income)
```{r}
summary_confint(lm(Avg_Hippocampal_volume_adjusted_T2_cm3~Natural_environment_300m_10pcnt + Age + Sex + BMI + Education + Income, CogsBrainCov.df2))
```

Full adjusted model (Covariates: Age, Sex, BMI, Education, Income, Average time spent outdoors, Physical activity minutes, Smoking status, Disability status, Townsend Deprivation Index, and Population Density)
```{r}
summary_confint(lm(Avg_Hippocampal_volume_adjusted_T2_cm3~Natural_environment_300m_10pcnt + Age + Sex + BMI + Education + Income + Average_time_outdoors + Summed_PA_min + Smoking + Disability + Deprivation_Index + Population_density, CogsBrainCov.df2))
```

### Model assumption checks
```{r}
#grey matter
model_diagnostics(lm(Grey_matter_volume_adjusted_T2_cm3~Natural_environment_300m_10pcnt + Age + Sex + BMI + Education + Income + Average_time_outdoors + Summed_PA_min + Smoking + Disability + Deprivation_Index + Population_density, CogsBrainCov.df2))

#white matter
model_diagnostics(lm(White_matter_volume_adjusted_T2_cm3~Natural_environment_300m_10pcnt + Age + Sex + BMI + Education + Income + Average_time_outdoors + Summed_PA_min + Smoking + Disability + Deprivation_Index + Population_density, CogsBrainCov.df2))

#HPC
model_diagnostics(lm(Avg_Hippocampal_volume_adjusted_T2_cm3~Natural_environment_300m_10pcnt + Age + Sex + BMI + Education + Income + Average_time_outdoors + Summed_PA_min + Smoking + Disability + Deprivation_Index + Population_density, CogsBrainCov.df2))
```


# 7) Moderation by sex for cognitive function [cognitive+brain+covariate subset]

## 7.1) 1000m natural environment buffer

### 7.1.1) Trails B-A (not significant)
```{r}
model7trail <- lm(Trails_B_A~Natural_environment_1000m_10pcnt*Sex + Age  + BMI + Education + Income + 
             Average_time_outdoors + 
             Summed_PA_min + Smoking + Disability + Deprivation_Index + Population_density, CogsBrainCov.df2)
summary(model7trail)
confint(model7trail,"Natural_environment_1000m_10pcnt:Sex") #95% confidence interval
```

### Model assumption checks
```{r}
#model_diagnostics(model7tmt)
```

### 7.1.2) Symbol Digit Substitution (not significant)
```{r}
model7sds <- lm(Symbol.Digit~Natural_environment_1000m_10pcnt*Sex + Age + BMI + Education + Income + 
             Average_time_outdoors + 
             Summed_PA_min + Smoking + Disability + Deprivation_Index + Population_density, CogsBrainCov.df2)
summary(model7sds)
confint(model7sds,"Natural_environment_1000m_10pcnt:Sex") #95% confidence interval
```

### Model assumption checks
```{r}
#model_diagnostics(model7sds)
```

## 7.2) 300m nature exposure buffer

### 7.2.1) Moderation by sex: association of natural environment with TMT (B-A) (*significant)

```{r}
summary_confint(lm(Trails_B_A~Natural_environment_300m_10pcnt*Sex + Age  + BMI + Education + Income + Average_time_outdoors + Summed_PA_min + Smoking + Disability + Deprivation_Index + Population_density, CogsBrainCov.df2))
```

TMT B-A: Sex-stratified subanalysis
```{r}
#stratify sample by sex
CogsBrainCov.F <- subset(CogsBrainCov.df2,Sex==0.5)
CogsBrainCov.M <- subset(CogsBrainCov.df2,Sex==-0.5)

#repeat regression within each Sex (ensuring to remove Sex as covariate)
#Males
lmM2 <- lm(Trails_B_A ~ Natural_environment_300m_10pcnt + Summed_PA_min + Age + BMI + Education + Income + Average_time_outdoors + Smoking + Disability + Deprivation_Index + Population_density, CogsBrainCov.M)
summary_confint(lmM2)

#Females
lmF2 <- lm(Trails_B_A ~ Natural_environment_300m_10pcnt + Summed_PA_min + Age + BMI + Education + Income + Average_time_outdoors + Smoking + Disability + Deprivation_Index + Population_density, CogsBrainCov.F)
summary_confint(lmF2)

#obtain beta coefficient and standard error for males and females
F.beta2 <- summary(lmF2)$coefficients["Natural_environment_300m_10pcnt","Estimate"]
F.se2 <- summary(lmF2)$coefficients["Natural_environment_300m_10pcnt","Std. Error"]

M.beta2 <- summary(lmM2)$coefficients["Natural_environment_300m_10pcnt","Estimate"]
M.se2 <- summary(lmM2)$coefficients["Natural_environment_300m_10pcnt","Std. Error"]

#calculate Z-score
Z.nat.tmt <- (M.beta2 - F.beta2) / sqrt(M.se2^2 + F.se2^2)
p.nat.tmt <- 1 - pnorm(Z.nat.tmt)

#print Z-score and p-value
Z.nat.tmt
p.nat.tmt
```




### Model assumption checks
```{r}
#model_diagnostics(lm(Trails_B_A~Natural_environment_300m_10pcnt*Sex + Age  + BMI + Education + Income + Average_time_outdoors + Summed_PA_min + Smoking + Disability + Deprivation_Index + Population_density, CogsBrainCov.df2))
```

### 7.2.2) Moderation by sex: association of natural environment with Symbol Digit Substitution (not significant)

```{r}
summary_confint(lm(Symbol.Digit~Natural_environment_300m_10pcnt*Sex + Age + BMI + Education + Income + 
             Average_time_outdoors + 
             Summed_PA_min + Smoking + Disability + Deprivation_Index + Population_density, CogsBrainCov.df2))
```

### Model assumption checks
```{r}
#model_diagnostics(lm(Symbol.Digit~Natural_environment_300m_10pcnt*Sex + Age  + BMI + Education + Income + Average_time_outdoors + Summed_PA_min + Smoking + Disability + Deprivation_Index + Population_density, CogsBrainCov.df2))
```


# 8) Moderation by sex for for brain structure [cognitive+brain+covariate subset]

## 8.1) 1000m natural environment buffer

### 8.1.1) Grey Matter (SIGNIFICANT)
```{r}
model8gm <- lm(Grey_matter_volume_adjusted_T2_cm3~Natural_environment_1000m_10pcnt*Sex + Age  + BMI + Education + Income + 
             Average_time_outdoors + 
             Summed_PA_min + Smoking + Disability + Deprivation_Index + Population_density, CogsBrainCov.df2)
summary(model8gm)
confint(model8gm,"Natural_environment_1000m_10pcnt:Sex") #95% confidence interval
```

Grey Matter Graph (no error bars)
```{r}

#maybe replace this with lmM (2 chunks below), depending on the purpose of scaling
Males.reg.mean<-lm(Grey_matter_volume_adjusted_T2_cm3~Natural_environment_pcnt_1000m + Age  + BMI + Education + Income + scale(Average_time_outdoors, TRUE) +
                        scale(Summed_PA_min, TRUE) +  Smoking + Disability + scale(Deprivation_Index, TRUE) + Population_density, CogsBrainCov.M)

Males.b0 <- round(Males.reg.mean$coeff[1], digits = 3)
Males.slope <- round(Males.reg.mean$coeff[2], digits = 3)
Males.ci <- predict(Males.reg.mean, interval = "confidence", level = 0.95)

Females.reg.mean<-lm(Grey_matter_volume_adjusted_T2_cm3~Natural_environment_pcnt_1000m + Age  + BMI + Education + Income + scale(Average_time_outdoors, TRUE) +
                     scale(Summed_PA_min, TRUE) +  Smoking + Disability + scale(Deprivation_Index, TRUE) + Population_density, CogsBrainCov.F)

Females.b0 <- round(Females.reg.mean$coeff[1], digits = 3)
Females.slope <- round(Females.reg.mean$coeff[2], digits = 3)
Females.ci <- predict(Females.reg.mean, interval = "confidence", level = 0.95)

ggplot() +
  #add data points but make colour white so invisible
  #(deleting the points also prevents regression lines from being visible)
  geom_point(data = CogsBrainCov.df2, aes(x = Natural_environment_pcnt_1000m, y = Grey_matter_volume_adjusted_T2_cm3), shape = 20, size = 0, color = "White") +
  #add regression lines for each sex
  geom_abline(aes(intercept= Males.b0, slope = Males.slope, color= "Males"), size = 1) + 
  geom_abline(aes(intercept = Females.b0, slope = Females.slope, color= "Females"), size = 1) +
  #add confidence intervals around regression lines for each sex
  #geom_ribbon(aes(x = CogsBrainCov.M$Natural_environment_pcnt_1000m, ymin = Males.ci[, "lwr"], ymax = Males.ci[,"upr"]), alpha = 0.3) +
  #geom_ribbon(aes(x = CogsBrainCov.F$Natural_environment_pcnt_1000m, ymin = Females.ci[, "lwr"], ymax = Females.ci[,"upr"]), alpha = 0.3) +
  #set plot labels, aesthetics, and legend
  guides(color = guide_legend((title = "Sex"))) + xlab("Natural Environment (%)") +
  ylab("Grey Matter Volume (cm^3)") + theme_bw() + theme(panel.border = element_blank(), panel.grid.major = element_blank(),
                                                    panel.grid.minor = element_blank(), axis.line = element_line(colour = "black")) +
 coord_cartesian(xlim=c(0,100),ylim=c(780,820))
```

Alternative Grey Matter Graph: Attempt to add error bands to regression lines
```{r}
# Ensure Education is a factor
CogsBrainCov.M$Education <- as.factor(CogsBrainCov.M$Education)
CogsBrainCov.F$Education <- as.factor(CogsBrainCov.F$Education)

# Ensure Age and BMI are numeric
CogsBrainCov.M$Age <- as.numeric(CogsBrainCov.M$Age)
CogsBrainCov.M$BMI <- as.numeric(CogsBrainCov.M$BMI)
CogsBrainCov.F$Age <- as.numeric(CogsBrainCov.F$Age)
CogsBrainCov.F$BMI <- as.numeric(CogsBrainCov.F$BMI)

# Fit the linear regression model
Males.reg.mean <- lm(Grey_matter_volume_adjusted_T2_cm3 ~ Natural_environment_pcnt_1000m + Age + BMI + Education + Income + scale(Average_time_outdoors) + scale(Summed_PA_min) +  Smoking + Disability + scale(Deprivation_Index) + Population_density, data = CogsBrainCov.M)
Females.reg.mean <- lm(Grey_matter_volume_adjusted_T2_cm3 ~ Natural_environment_pcnt_1000m + Age + BMI + Education + Income + scale(Average_time_outdoors) + scale(Summed_PA_min) +  Smoking + Disability + scale(Deprivation_Index) + Population_density, data = CogsBrainCov.F)

# Create new data frame for prediction
new_data_M <- data.frame(
    Natural_environment_pcnt_1000m = c(1:100),
    Age = rep(0, 100),
    BMI = rep(0, 100),
    Education = factor(rep(levels(CogsBrainCov.M$Education)[1], 100), levels = levels(CogsBrainCov.M$Education)),
    Income = rep(mean(CogsBrainCov.M$Income, na.rm = TRUE), 100),
    Average_time_outdoors = rep(mean(CogsBrainCov.M$Average_time_outdoors, na.rm = TRUE), 100),
    Summed_PA_min = rep(mean(CogsBrainCov.M$Summed_PA_min, 100), 100),
    Smoking = rep(mean(CogsBrainCov.M$Smoking, na.rm = TRUE), 100),
    Disability = rep(mean(CogsBrainCov.M$Disability, na.rm = TRUE), 100),
    Deprivation_Index = rep(mean(CogsBrainCov.M$Deprivation_Index, na.rm = TRUE), 100),
    Population_density = rep(mean(CogsBrainCov.M$Population_density, na.rm = TRUE), 100)
)
new_data_F <- data.frame(
    Natural_environment_pcnt_1000m = c(1:100),
    Age = rep(0, 100),
    BMI = rep(0, 100),
    Education = factor(rep(levels(CogsBrainCov.F$Education)[1], 100), levels = levels(CogsBrainCov.F$Education)),
    Income = rep(mean(CogsBrainCov.F$Income, na.rm = TRUE), 100),
    Average_time_outdoors = rep(mean(CogsBrainCov.F$Average_time_outdoors, na.rm = TRUE), 100),
    Summed_PA_min = rep(mean(CogsBrainCov.F$Summed_PA_min, 100), 100),
    Smoking = rep(mean(CogsBrainCov.F$Smoking, na.rm = TRUE), 100),
    Disability = rep(mean(CogsBrainCov.F$Disability, na.rm = TRUE), 100),
    Deprivation_Index = rep(mean(CogsBrainCov.F$Deprivation_Index, na.rm = TRUE), 100),
    Population_density = rep(mean(CogsBrainCov.F$Population_density, na.rm = TRUE), 100)
)

# Ensure that all variables are of the correct type
new_data_M$Age <- as.numeric(new_data_M$Age)
new_data_M$BMI <- as.numeric(new_data_M$BMI)
new_data_F$Age <- as.numeric(new_data_F$Age)
new_data_F$BMI <- as.numeric(new_data_F$BMI)

# Get predictions and confidence intervals
predictions_M <- predict(Males.reg.mean, newdata = new_data_M, interval = "confidence", level = 0.95)
predictions_F <- predict(Females.reg.mean, newdata = new_data_F, interval = "confidence", level = 0.95)

# Combine predictions with new_data
new_data_M <- cbind(new_data_M, predictions_M)
new_data_F <- cbind(new_data_F, predictions_F)

# Plot the data, regression line, and confidence intervals
ggplot() +
    geom_line(data = new_data_M, aes(x = Natural_environment_pcnt_1000m, y = fit, color = "Males")) +
    geom_line(data = new_data_F, aes(x = Natural_environment_pcnt_1000m, y = fit, color = "Females")) +
    geom_ribbon(data = new_data_M, aes(x = Natural_environment_pcnt_1000m, ymin = lwr, ymax = upr), alpha = 0.2, fill = "Green") +
    geom_ribbon(data = new_data_F, aes(x = Natural_environment_pcnt_1000m, ymin = lwr, ymax = upr), alpha = 0.2, fill = "Red") +
    guides(color = guide_legend(title = "Sex")) +
  labs(x = "Natural Environment (%)", y = "Grey Matter Volume (cm^3)", color =  "Sex") + 
    theme_classic() 
```

Alternative Grey Matter Graph: Partial Regression Plot
```{r}
#residualize GM volume and natural environment within Males
M.resid.gm <- resid(lm(Grey_matter_volume_adjusted_T2_cm3~Age + BMI + Education + Income +                          Average_time_outdoors + Summed_PA_min + Smoking + Disability + 
                         Deprivation_Index + Population_density, CogsBrainCov.M))
M.resid.nat <- resid(lm(Natural_environment_1000m_10pcnt~Age + BMI + Education + Income + 
                  Average_time_outdoors + Summed_PA_min + Smoking + Disability + 
                    Deprivation_Index + Population_density, CogsBrainCov.M))

#residualize GM volume and natural environment within Females
F.resid.gm <- resid(lm(Grey_matter_volume_adjusted_T2_cm3~Age + BMI + Education + Income +                          Average_time_outdoors + Summed_PA_min + Smoking + Disability + 
                         Deprivation_Index + Population_density, CogsBrainCov.F))
F.resid.nat <- resid(lm(Natural_environment_1000m_10pcnt~Age + BMI + Education + Income + 
                  Average_time_outdoors + Summed_PA_min + Smoking + Disability + 
                    Deprivation_Index + Population_density, CogsBrainCov.F))

#make data frame of MALE residuals for natural environment and GM volume
M.resid <- as.data.frame(cbind(M.resid.nat, M.resid.gm))
#do NOT add the brain and natural environment means to their respective residuals

#make data frame of FEMALE residuals for natural environment and GM volume
F.resid <- as.data.frame(cbind(F.resid.nat, F.resid.gm))
# do NOT add the FEMALE brain and cog means to their respective residuals

#Plot the partial regression, including confidence intervals
ggplot() +
  #add regression line for females
  geom_smooth(data = F.resid, aes(x = F.resid.nat*10, y = F.resid.gm, color = "Females"), method = "lm", se = TRUE, fill = "red", alpha = 0.1) +
  #add regression line for males
  geom_smooth(data = M.resid, aes(x = M.resid.nat*10, y = M.resid.gm, color = "Males"), method = "lm", se = TRUE, fill = "green", alpha = 0.1) +
  #set plot labels, aesthetics, and legend
  labs(x = "Natural Environment (%)", y = "Grey Matter Volume (cm^3)", color = "Sex") +
  #remove the fill colour from the legend symbols
  guides(color = guide_legend(override.aes = list(fill = NA))) +
  #xlim(0,100) +
  #ylim(785,795) +
  theme_classic()
```

Alternative Grey Matter Graph: Partial Regression Plot #2
```{r}
#residualize GM volume and natural environment within Males
M.resid.gm <- resid(lm(Grey_matter_volume_adjusted_T2_cm3~Age + BMI + Education + Income +                          Average_time_outdoors + Summed_PA_min + Smoking + Disability + 
                         Deprivation_Index + Population_density, CogsBrainCov.M))
M.resid.nat <- resid(lm(Natural_environment_1000m_10pcnt~Age + BMI + Education + Income + 
                  Average_time_outdoors + Summed_PA_min + Smoking + Disability + 
                         Deprivation_Index + Population_density, CogsBrainCov.M))

#residualize GM volume and natural environment within Females
F.resid.gm <- resid(lm(Grey_matter_volume_adjusted_T2_cm3~Age + BMI + Education + Income +                          Average_time_outdoors + Summed_PA_min + Smoking + Disability + 
                         Deprivation_Index + Population_density, CogsBrainCov.F))
F.resid.nat <- resid(lm(Natural_environment_1000m_10pcnt~Age + BMI + Education + Income + 
                  Average_time_outdoors + Summed_PA_min + Smoking + Disability + 
                    Deprivation_Index + Population_density, CogsBrainCov.F))

#make data frame of MALE residuals for natural environment and GM volume
M.resid <- as.data.frame(cbind(M.resid.nat, M.resid.gm))
#add the MALE brain and natural environment means to their respective residuals
M.resid$M.resid.gm <- M.resid$M.resid.gm + mean(CogsBrainCov.M$Grey_matter_volume_adjusted_T2_cm3)
M.resid$M.resid.nat <- M.resid$M.resid.nat*10 + mean(CogsBrainCov.M$Natural_environment_1000m_10pcnt)*10

#make data frame of FEMALE residuals for natural environment and GM volume
F.resid <- as.data.frame(cbind(F.resid.nat, F.resid.gm))
#add the FEMALE brain and cog means to their respective residuals
F.resid$F.resid.gm <- F.resid$F.resid.gm + mean(CogsBrainCov.F$Grey_matter_volume_adjusted_T2_cm3)
F.resid$F.resid.nat <- F.resid$F.resid.nat*10 + mean(CogsBrainCov.F$Natural_environment_1000m_10pcnt)*10

ggplot() +
  #add data points for females
  #geom_point(data = F.resid, aes(x = F.resid.nat, y = F.resid.gm), shape = 20, size = 0, color = "Red") +
  #add data points for males
  #geom_point(data = M.resid, aes(x = M.resid.nat, y = M.resid.gm), shape = 3, size = 0, color = "Green") +
  #add regression line for females
  geom_smooth(data = F.resid, aes(x = F.resid.nat, y = F.resid.gm, color = "Females"), method = "lm", se = TRUE, fill = "red", alpha = 0.1) +
  #add regression line for males
  geom_smooth(data = M.resid, aes(x = M.resid.nat, y = M.resid.gm, color = "Males"), method = "lm", se = TRUE, fill = "green", alpha = 0.2) +
  #set plot labels, aesthetics, and legend
  labs(x = "Natural Environment (%)", y = "Grey Matter Volume (cm^3)", color = "Sex") +
  #remove the fill colour from the legend symbols
  guides(color = guide_legend(override.aes = list(fill = NA))) +
  coord_cartesian(xlim=c(0,96), ylim=c(765,815)) +
  theme_classic()

#What is going wrong with these graphs? Why are the slopes so shallow?
#Based on the sex-stratified models (below), y-intercept should be 786 for Males and 813 for Females. Why the discrepancy????
```

Grey matter: Sex-stratified subanalysis
```{r}
#repeat regression within each Sex (ensuring to remove Sex as covariate)
#Males
lmM <- lm(Grey_matter_volume_adjusted_T2_cm3 ~ Natural_environment_1000m_10pcnt + Summed_PA_min + Age + BMI + Education + Income + Average_time_outdoors + Smoking + Disability + Deprivation_Index + Population_density, CogsBrainCov.M)
summary(lmM)
confint(lmM,"Natural_environment_1000m_10pcnt")

#Females
lmF <- lm(Grey_matter_volume_adjusted_T2_cm3 ~ Natural_environment_1000m_10pcnt + Summed_PA_min + Age + BMI + Education + Income + Average_time_outdoors + Smoking + Disability + Deprivation_Index + Population_density, CogsBrainCov.F)
summary(lmF)
confint(lmF,"Natural_environment_1000m_10pcnt")

#obtain beta coefficient and standard error for males and females
F.beta <- summary(lmF)$coefficients["Natural_environment_1000m_10pcnt","Estimate"]
F.se <- summary(lmF)$coefficients["Natural_environment_1000m_10pcnt","Std. Error"]

M.beta <- summary(lmM)$coefficients["Natural_environment_1000m_10pcnt","Estimate"]
M.se <- summary(lmM)$coefficients["Natural_environment_1000m_10pcnt","Std. Error"]

#calculate Z-score
Z.nat.gm <- (M.beta - F.beta) / sqrt(M.se^2 + F.se^2)
p.nat.gm <- 1 - pnorm(Z.nat.gm)

#print Z-score and p-value
Z.nat.gm
p.nat.gm
```


### 8.1.2) White Matter (not significant)
```{r}
model8wm <- lm(White_matter_volume_adjusted_T2_cm3~Natural_environment_1000m_10pcnt*Sex + Age + BMI + Education + Income + 
             Average_time_outdoors + 
             Summed_PA_min + Smoking + Disability + Deprivation_Index + Population_density, CogsBrainCov.df2)
summary(model8wm)
confint(model8wm,"Natural_environment_1000m_10pcnt:Sex") #95% confidence interval
```


### 8.1.3) Hippocampal volume (not significant)
```{r}
model8hpc <- lm(Avg_Hippocampal_volume_adjusted_T2_cm3~Natural_environment_1000m_10pcnt*Sex + Age + BMI + Education + Income + 
             Average_time_outdoors + 
             Summed_PA_min + Smoking + Disability + Deprivation_Index + Population_density, CogsBrainCov.df2)
summary(model8hpc)
confint(model8hpc,"Natural_environment_1000m_10pcnt:Sex") #95% confidence interval
```

## Model assumption checks
```{r}
model_diagnostics(model8gm)
  # Residuals vs. Fitted plot suggests approximately linear
  # Normal Q-Q plot suggests slight non-normal skew
  # Residuals vs. Leverage plot: no concerning outliers
  # Histogram of residuals: approximately normal
  # Durbin-Watson test near 2 suggests independent residuals
  # Skewness, Kurtosis, and Heteroskedasticity tests suggest non-normal but homoskedastic

model_diagnostics(model8wm)
  # Residuals vs. Fitted plot suggests approximately linear, some heteroskedasticity, 1 outlier
  # Normal Q-Q plot suggests non-normal skew
  # Residuals vs. Leverage plot: no concerning outliers
  # Histogram of residuals: approximately normal
  # Durbin-Watson test near 2 suggests independent residuals
  # Skewness, Kurtosis, and Heteroskedasticity tests suggest non-normal but homoskedastic

model_diagnostics(model8hpc, bin_width=0.1)
  # Residuals vs. Fitted plot suggests approximately linear, some heteroskedasticity, 1 outlier
  # Normal Q-Q plot suggests non-normal kurtosis
  # Residuals vs. Leverage plot: a few outliers
  # Histogram of residuals: approximately normal
  # Durbin-Watson test near 2 suggests independent residuals
  # Skewness, Kurtosis, and Heteroskedasticity tests suggest non-normal but homoskedastic
```


## 8.2) 300m natural environment buffer

### 8.2.1) Moderation by sex: Grey Matter (significant)
```{r}
summary_confint(lm(Grey_matter_volume_adjusted_T2_cm3~Natural_environment_300m_10pcnt*Sex + Age  + BMI + Education + Income + 
             Average_time_outdoors + 
             Summed_PA_min + Smoking + Disability + Deprivation_Index + Population_density, CogsBrainCov.df2))
```

Grey matter: Sex-stratified subanalysis
```{r}
#repeat regression within each Sex (ensuring to remove Sex as covariate)
#Males
lmM <- lm(Grey_matter_volume_adjusted_T2_cm3 ~ Natural_environment_300m_10pcnt + Summed_PA_min + Age + BMI + Education + Income + Average_time_outdoors + Smoking + Disability + Deprivation_Index + Population_density, CogsBrainCov.M)
summary(lmM)
confint(lmM,"Natural_environment_300m_10pcnt")

#Females
lmF <- lm(Grey_matter_volume_adjusted_T2_cm3 ~ Natural_environment_300m_10pcnt + Summed_PA_min + Age + BMI + Education + Income + Average_time_outdoors + Smoking + Disability + Deprivation_Index + Population_density, CogsBrainCov.F)
summary(lmF)
confint(lmF,"Natural_environment_300m_10pcnt")

#obtain beta coefficient and standard error for males and females
F.beta <- summary(lmF)$coefficients["Natural_environment_300m_10pcnt","Estimate"]
F.se <- summary(lmF)$coefficients["Natural_environment_300m_10pcnt","Std. Error"]

M.beta <- summary(lmM)$coefficients["Natural_environment_300m_10pcnt","Estimate"]
M.se <- summary(lmM)$coefficients["Natural_environment_300m_10pcnt","Std. Error"]

#calculate Z-score
Z.nat.gm <- (M.beta - F.beta) / sqrt(M.se^2 + F.se^2)
p.nat.gm <- 1 - pnorm(Z.nat.gm)

#print Z-score and p-value
Z.nat.gm
p.nat.gm
```


### 8.2.2) Moderation by sex: White Matter (not significant)
```{r}
summary_confint(lm(White_matter_volume_adjusted_T2_cm3~Natural_environment_300m_10pcnt*Sex + Age  + BMI + Education + Income + 
             Average_time_outdoors + 
             Summed_PA_min + Smoking + Disability + Deprivation_Index + Population_density, CogsBrainCov.df2))
```

### 8.2.3) Moderation by sex: Average hippocampal volume (not significant)
```{r}
summary_confint(lm(Avg_Hippocampal_volume_adjusted_T2_cm3~Natural_environment_300m_10pcnt*Sex + Age  + BMI + Education + Income + 
             Average_time_outdoors + 
             Summed_PA_min + Smoking + Disability + Deprivation_Index + Population_density, CogsBrainCov.df2))
```


# 9) Sensitivity analyses for cognitive function

## 9.1) Greenspace percentage instead of natural environment percentage

### 9.1.1) 1000m greenspace buffer

#### Association of greenspace (buffered to 1000m) with Trail-making Test (B-A) (not significant)
Crude model (Covariates: Age, Sex, BMI, Education, and Income)
```{r}
summary_confint(lm(Trails_B_A~Greenspace_1000m_10pcnt + Age + Sex + BMI + Education + Income, CogsBrainCov.df2))
```

Full adjusted model (Covariates: Age, Sex, BMI, Education, Income, Average time spent outdoors, Physical activity minutes, Smoking status, Disability status, Townsend Deprivation Index, and Population Density)
```{r}
summary_confint(lm(Trails_B_A~Greenspace_1000m_10pcnt + Age + Sex + BMI + Education + Income + 
             Average_time_outdoors + 
             Summed_PA_min + Smoking + Disability + Deprivation_Index + Population_density, CogsBrainCov.df2))
```


#### Association of greenspace (buffered to 1000m) with Symbol Digit Substitution (significant)

Crude model (Covariates: Age, Sex, BMI, Education, and Income)
```{r}
summary_confint(lm(Symbol.Digit~Greenspace_1000m_10pcnt + Age + Sex + BMI + Education + Income, CogsBrainCov.df2))
```

Full adjusted model (Covariates: Age, Sex, BMI, Education, Income, Average time spent outdoors, Physical activity minutes, Smoking status, Disability status, Townsend Deprivation Index, and Population Density)
```{r}
summary_confint(lm(Symbol.Digit~Greenspace_1000m_10pcnt + Age + Sex + BMI + Education + Income + 
             Average_time_outdoors + 
             Summed_PA_min + Smoking + Disability + Deprivation_Index + Population_density, CogsBrainCov.df2))
```


#### Moderation by sex: association of greenspace (buffered to 1000m) with TMT (B-A) (not significant)

```{r}
summary_confint(lm(Trails_B_A~Greenspace_1000m_10pcnt*Sex + Age  + BMI + Education + Income + 
             Average_time_outdoors + 
             Summed_PA_min + Smoking + Disability + Deprivation_Index + Population_density, CogsBrainCov.df2))
```

#### Moderation by sex: association of greenspace (buffered to 1000m) with Symbol Digit Substitution (not significant)
```{r}
summary_confint(lm(Symbol.Digit~Greenspace_1000m_10pcnt*Sex + Age + BMI + Education + Income + 
             Average_time_outdoors + 
             Summed_PA_min + Smoking + Disability + Deprivation_Index + Population_density, CogsBrainCov.df2))
```

### 9.1.2) 300m greenspace buffer

#### Association of greenspace (buffered to 300m) with Trail-making Test (B-A) (not significant)
```{r}
summary_confint(lm(Trails_B_A~Greenspace_300m_10pcnt + Age + Sex + BMI + Education + Income + 
             Average_time_outdoors + 
             Summed_PA_min + Smoking + Disability + Deprivation_Index + Population_density, CogsBrainCov.df2))
```

#### Association of greenspace (buffered to 300m) with Symbol Digit Substitution (not significant)

Full adjusted model (Covariates: Age, Sex, BMI, Education, Income, Average time spent outdoors, Physical activity minutes, Smoking status, Disability status, Townsend Deprivation Index, and Population Density)
```{r}
summary_confint(lm(Symbol.Digit~Greenspace_300m_10pcnt + Age + Sex + BMI + Education + Income + 
             Average_time_outdoors + 
             Summed_PA_min + Smoking + Disability + Deprivation_Index + Population_density, CogsBrainCov.df2))
```

#### Moderation by sex: association of greenspace (buffered to 300m) with TMT (B-A) (not significant)
```{r}
summary_confint(lm(Trails_B_A~Greenspace_300m_10pcnt*Sex + Age  + BMI + Education + Income + 
             Average_time_outdoors + 
             Summed_PA_min + Smoking + Disability + Deprivation_Index + Population_density, CogsBrainCov.df2))
```

#### Moderation by sex: association of greenspace (buffered to 300m) with Symbol Digit Substitution (not significant)
```{r}
summary_confint(lm(Symbol.Digit~Greenspace_300m_10pcnt*Sex + Age + BMI + Education + Income + 
             Average_time_outdoors + 
             Summed_PA_min + Smoking + Disability + Deprivation_Index + Population_density, CogsBrainCov.df2))
```


## 9.2) Restrict sample to participants with no change in home location between baseline (20062010) and online cognitive assessment (2014-2015, individual-specific dates)

Required UKB variables:
* 699 - Length of time at current address - Instances 0 and 2 (unit: years)
* 53 - Date of attending assessment centre - Instances 0 and 2 (format: YYYY-MM-DD)
Approach: Include if var 699 (Instance) >= Difference in 53 (Instance 2 minus Instance 0)
Note: This is a conservative estimate of the subset of participants with no change in home address between baseline and 2014-2015. We were unable to use the actual dates of the online cognitive assessment (var 20136 and 20137; below) because variable 699 was not collected at the same time as the online cognitive assessment; the nearest time at which variable 699 was collected is Instance 2.
* 20136 - When trail making test completed
* 20137 - When symbol digit substitution test completed

```{r}
#convert dates to date data type
CogsBrainCov.df2$Date_attending_centre_0 <- as.Date(CogsBrainCov.df2$Date.attending.centre.0)
CogsBrainCov.df2$Date_attending_centre_2 <- as.Date(CogsBrainCov.df2$Date.attending.centre.2)
CogsBrainCov.df2$Date_attending_centre_3 <- as.Date(CogsBrainCov.df2$Date.attending.centre.3)

#include only observations with no missing data for Var 699 and Var 53
HomeChangeTest_T2 <- subset(CogsBrainCov.df2, !is.na(Date_attending_centre_0) & !is.na(Date_attending_centre_2) & !is.na(Time_at_current_address_T2))
length(HomeChangeTest_T2$ID)
#include only observations with no change in address between T2 and T0
NoHomeChange_T2 <- subset(HomeChangeTest_T2,Time_at_current_address_T2 >= (round(as.numeric(Date_attending_centre_2 - Date_attending_centre_0) / 365, digits = 0)))
  #more stringent version, where time difference between T2 and T0 is not rounded:
  # NoHomeChange_T2 <- subset(HomeChangeTest_T2,Time_at_current_address_T2 >= (as.numeric(Date_attending_centre_2 - Date_attending_centre_0) / 365, digits = 0))
length(NoHomeChange_T2$ID)
```

### Compare the analytical cognitive+brain+covariate data set to the restricted data set with no change in home location
```{r eval= FALSE}
#use CogsBrainCov.df because variables (such as age, BMI, natural environment) are not yet re-scaled

#convert dates to date data type
CogsBrainCov.df$Date_attending_centre_0 <- as.Date(CogsBrainCov.df$Date.attending.centre.0)
CogsBrainCov.df$Date_attending_centre_2 <- as.Date(CogsBrainCov.df$Date.attending.centre.2)
CogsBrainCov.df$Date_attending_centre_3 <- as.Date(CogsBrainCov.df$Date.attending.centre.3)

#include only observations with no missing data for Var 699 and Var 53
HomeChangeTest_T2_unscaled <- subset(CogsBrainCov.df, !is.na(Date_attending_centre_0) & !is.na(Date_attending_centre_2) & !is.na(Time_at_current_address_T2))

#include only observations with no change in address between T2 and T0
NoHomeChange_T2_unscaled <- subset(HomeChangeTest_T2_unscaled,Time_at_current_address_T2 >= (round(as.numeric(Date_attending_centre_2 - Date_attending_centre_0) / 365, digits = 0)))

NoHomeChange_T2_unscaled$included<-"Participants with no change in home location"

AnalyticalVsNoHomeChange<-rbind(CogsBrainCov.df, NoHomeChange_T2_unscaled)

vars2<- subset(AnalyticalVsNoHomeChange, select = c(2:44, 49))

vars2 %>% tbl_summary(
  statistic = list(all_continuous() ~ "{mean} ({sd})",
                   all_categorical() ~ "{n} ({p}%)"),
  digits = all_continuous() ~ 2,
  by = "included"
)
```

Table for restricted sample with no change in home location, stratified by sex
```{r eval= FALSE}
library(dplyr)

tbl_summary(
  select(NoHomeChange_T2_unscaled, c(2:43)),
  by = "Sex", 
  statistic = list(all_continuous() ~ "{mean} ({sd})",
                   all_categorical() ~ "{n} ({p}%)"),
  digits = all_continuous() ~ 2
) %>%
  add_p(test = list(all_continuous() ~ "t.test",
                   all_categorical() ~ "chisq.test"))
```

### 9.2.1) 1000m natural environment buffer

#### Association of natural environment with Trail-making Test (B-A) (not significant)
Crude model (Covariates: Age, Sex, BMI, Education, and Income)
```{r}
summary_confint(lm(Trails_B_A~Natural_environment_1000m_10pcnt + Age + Sex + BMI + Education + Income, NoHomeChange_T2))
```

Full adjusted model (Covariates: Age, Sex, BMI, Education, Income, Average time spent outdoors, Physical activity minutes, Smoking status, Disability status, Townsend Deprivation Index, and Population Density)
```{r}
summary_confint(lm(Trails_B_A~Natural_environment_1000m_10pcnt + Age + Sex + BMI + Education + Income + 
             Average_time_outdoors + 
             Summed_PA_min + Smoking + Disability + Deprivation_Index + Population_density, NoHomeChange_T2))
```


#### Association of natural environment with Symbol Digit Substitution (significant)

Crude model (Covariates: Age, Sex, BMI, Education, and Income)
```{r}
summary_confint(lm(Symbol.Digit~Natural_environment_1000m_10pcnt + Age + Sex + BMI + Education + Income, NoHomeChange_T2))
```

Full adjusted model (Covariates: Age, Sex, BMI, Education, Income, Average time spent outdoors, Physical activity minutes, Smoking status, Disability status, Townsend Deprivation Index, and Population Density)
```{r}
summary_confint(lm(Symbol.Digit~Natural_environment_1000m_10pcnt + Age + Sex + BMI + Education + Income + 
             Average_time_outdoors + 
             Summed_PA_min + Smoking + Disability + Deprivation_Index + Population_density, NoHomeChange_T2))
```

#### Moderation by sex: association of natural environment with TMT (B-A) (not significant)

```{r}
summary_confint(lm(Trails_B_A~Natural_environment_1000m_10pcnt*Sex + Age  + BMI + Education + Income + 
             Average_time_outdoors + 
             Summed_PA_min + Smoking + Disability + Deprivation_Index + Population_density, NoHomeChange_T2))
```

#### Moderation by sex: association of natural environment with Symbol Digit Substitution (not significant)
```{r}
summary_confint(lm(Symbol.Digit~Natural_environment_1000m_10pcnt*Sex + Age + BMI + Education + Income + 
             Average_time_outdoors + 
             Summed_PA_min + Smoking + Disability + Deprivation_Index + Population_density, NoHomeChange_T2))
```

### 9.2.2) 300m natural environment buffer


#### Association of natural environment (300m buffer) with Trail-making Test (B-A)

Full adjusted model (Covariates: Age, Sex, BMI, Education, Income, Average time spent outdoors, Physical activity minutes, Smoking status, Disability status, Townsend Deprivation Index, and Population Density)
```{r}
summary_confint(lm(Trails_B_A~Natural_environment_300m_10pcnt + Age + Sex + BMI + Education + Income + 
             Average_time_outdoors + 
             Summed_PA_min + Smoking + Disability + Deprivation_Index + Population_density, NoHomeChange_T2))
```

#### Association of natural environment (300m buffer) with Symbol Digit Substitution

Full adjusted model (Covariates: Age, Sex, BMI, Education, Income, Average time spent outdoors, Physical activity minutes, Smoking status, Disability status, Townsend Deprivation Index, and Population Density)
```{r}
summary_confint(lm(Symbol.Digit~Natural_environment_300m_10pcnt + Age + Sex + BMI + Education + Income + 
             Average_time_outdoors + 
             Summed_PA_min + Smoking + Disability + Deprivation_Index + Population_density, NoHomeChange_T2))
```

#### Moderation by sex: association of natural environment (300m buffer) with TMT (B-A)
```{r}
summary_confint(lm(Trails_B_A~Natural_environment_300m_10pcnt*Sex + Age  + BMI + Education + Income + 
             Average_time_outdoors + 
             Summed_PA_min + Smoking + Disability + Deprivation_Index + Population_density, NoHomeChange_T2))
```

#### Moderation by sex: association of natural environment (300m buffer) with Symbol Digit Substitution
```{r}
summary_confint(lm(Symbol.Digit~Natural_environment_300m_10pcnt*Sex + Age + BMI + Education + Income + 
             Average_time_outdoors + 
             Summed_PA_min + Smoking + Disability + Deprivation_Index + Population_density, NoHomeChange_T2))
```


# 10) Sensitivity analyses for brain volumes

## 10.1) Greenspace percentage instead of natural environment percentage

### 10.1.1) 1000m greenspace buffer

#### Association of greenspace (buffered to 1000m) with grey matter volume (significant)

Crude model (Covariates: Age, Sex, BMI, Education, and Income)
```{r}
summary_confint(lm(Grey_matter_volume_adjusted_T2_cm3~Greenspace_1000m_10pcnt + Age + Sex + BMI + Education + Income, CogsBrainCov.df2))
```

Full adjusted model (Covariates: Age, Sex, BMI, Education, Income, Average time spent outdoors, Physical activity minutes, Smoking status, Disability status, Townsend Deprivation Index, and Population Density)
```{r}
summary_confint(lm(Grey_matter_volume_adjusted_T2_cm3~Greenspace_1000m_10pcnt + Age + Sex + BMI + Education + Income + Average_time_outdoors + Summed_PA_min + Smoking + Disability + Deprivation_Index + Population_density, CogsBrainCov.df2))
```

#### Association of greenspace (buffered to 1000m) with white matter volume (significant)

Crude model (Covariates: Age, Sex, BMI, Education, and Income)
```{r}
summary_confint(lm(White_matter_volume_adjusted_T2_cm3~Greenspace_1000m_10pcnt + Age + Sex + BMI + Education + Income, CogsBrainCov.df2))
```

Full adjusted model (Covariates: Age, Sex, BMI, Education, Income, Average time spent outdoors, Physical activity minutes, Smoking status, Disability status, Townsend Deprivation Index, and Population Density)
```{r}
summary_confint(lm(White_matter_volume_adjusted_T2_cm3~Greenspace_1000m_10pcnt + Age + Sex + BMI + Education + Income + Average_time_outdoors + Summed_PA_min + Smoking + Disability + Deprivation_Index + Population_density, CogsBrainCov.df2))
```


#### Association of greenspace (buffered to 1000m) with hippocampal volume (not significant)

Crude model (Covariates: Age, Sex, BMI, Education, and Income)
```{r}
summary_confint(lm(Avg_Hippocampal_volume_adjusted_T2_cm3~Greenspace_1000m_10pcnt + Age + Sex + BMI + Education + Income, CogsBrainCov.df2))
```

Full adjusted model (Covariates: Age, Sex, BMI, Education, Income, Average time spent outdoors, Physical activity minutes, Smoking status, Disability status, Townsend Deprivation Index, and Population Density)
```{r}
summary_confint(lm(Avg_Hippocampal_volume_adjusted_T2_cm3~Greenspace_1000m_10pcnt + Age + Sex + BMI + Education + Income + Average_time_outdoors + Summed_PA_min + Smoking + Disability + Deprivation_Index + Population_density, CogsBrainCov.df2))
```

#### Moderation by sex: Grey Matter (significant)
```{r}
summary_confint(lm(Grey_matter_volume_adjusted_T2_cm3~Greenspace_1000m_10pcnt*Sex + Age  + BMI + Education + Income + 
             Average_time_outdoors + 
             Summed_PA_min + Smoking + Disability + Deprivation_Index + Population_density, CogsBrainCov.df2))
```

#### Moderation by sex: White Matter (not significant)
```{r}
summary_confint(lm(White_matter_volume_adjusted_T2_cm3~Greenspace_1000m_10pcnt*Sex + Age  + BMI + Education + Income + 
             Average_time_outdoors + 
             Summed_PA_min + Smoking + Disability + Deprivation_Index + Population_density, CogsBrainCov.df2))
```

#### Moderation by sex: Average hippocampal volume (not significant)
```{r}
summary_confint(lm(Avg_Hippocampal_volume_adjusted_T2_cm3~Greenspace_1000m_10pcnt*Sex + Age  + BMI + Education + Income + 
             Average_time_outdoors + 
             Summed_PA_min + Smoking + Disability + Deprivation_Index + Population_density, CogsBrainCov.df2))
```


### 10.1.2) 300m greenspace buffer

#### Association of greenspace (buffered to 300m) with grey matter volume (significant)

Full adjusted model (Covariates: Age, Sex, BMI, Education, Income, Average time spent outdoors, Physical activity minutes, Smoking status, Disability status, Townsend Deprivation Index, and Population Density)
```{r}
summary_confint(lm(Grey_matter_volume_adjusted_T2_cm3~Greenspace_300m_10pcnt + Age + Sex + BMI + Education + Income + Average_time_outdoors + Summed_PA_min + Smoking + Disability + Deprivation_Index + Population_density, CogsBrainCov.df2))
```

#### Association of greenspace (buffered to 300m) with white matter volume (significant)

Full adjusted model (Covariates: Age, Sex, BMI, Education, Income, Average time spent outdoors, Physical activity minutes, Smoking status, Disability status, Townsend Deprivation Index, and Population Density)
```{r}
summary_confint(lm(White_matter_volume_adjusted_T2_cm3~Greenspace_300m_10pcnt + Age + Sex + BMI + Education + Income + Average_time_outdoors + Summed_PA_min + Smoking + Disability + Deprivation_Index + Population_density, CogsBrainCov.df2))
```

#### Association of greenspace (buffered to 300m) with hippocampal volume (not significant)

Full adjusted model (Covariates: Age, Sex, BMI, Education, Income, Average time spent outdoors, Physical activity minutes, Smoking status, Disability status, Townsend Deprivation Index, and Population Density)
```{r}
summary_confint(lm(Avg_Hippocampal_volume_adjusted_T2_cm3~Greenspace_300m_10pcnt + Age + Sex + BMI + Education + Income + Average_time_outdoors + Summed_PA_min + Smoking + Disability + Deprivation_Index + Population_density, CogsBrainCov.df2))
```

#### Moderation by sex: Grey Matter [Greenspace AND 300m buffer] (significant)
```{r}
summary_confint(lm(Grey_matter_volume_adjusted_T2_cm3~Greenspace_300m_10pcnt*Sex + Age  + BMI + Education + Income + 
             Average_time_outdoors + 
             Summed_PA_min + Smoking + Disability + Deprivation_Index + Population_density, NoHomeChange_T2))
```

#### Moderation by sex: White Matter [Greenspace AND 300m buffer] (not significant)
```{r}
summary_confint(lm(White_matter_volume_adjusted_T2_cm3~Greenspace_300m_10pcnt*Sex + Age  + BMI + Education + Income + 
             Average_time_outdoors + 
             Summed_PA_min + Smoking + Disability + Deprivation_Index + Population_density, CogsBrainCov.df2))
```

#### Moderation by sex: Average hippocampal volume [Greenspace AND 300m buffer] (not significant) (not significant)
```{r}
summary_confint(lm(Avg_Hippocampal_volume_adjusted_T2_cm3~Greenspace_300m_10pcnt*Sex + Age  + BMI + Education + Income + 
             Average_time_outdoors + 
             Summed_PA_min + Smoking + Disability + Deprivation_Index + Population_density, CogsBrainCov.df2))
```

## 10.2) Restrict sample to participants with no change in home location between baseline (20062010) and first MRI assessment (2014+)

Required UKB variables:
* 699 - Length of time at current address - Instances 2 (measured in units of whole years, except if answer was < 1 year, in which case the value assigned was 0.5 years)
* 53 - Date of attending assessment centre - Instances 0 and 2 (format: YYYY-MM-DD)
Approach: Include if var 699 (Instance 2) >= Difference in 53 (Instance 2 minus Instance 0)
Note: This subset, "NoHomeChange_T2", was generated in Section 10.3.

### 10.2.1) 1000m natural environment buffer

#### Association of natural environment with grey matter volume (significant)

Crude model (Covariates: Age, Sex, BMI, Education, and Income)
```{r}
summary_confint(lm(Grey_matter_volume_adjusted_T2_cm3~Natural_environment_1000m_10pcnt + Age + Sex + BMI + Education + Income, NoHomeChange_T2))
```

Full adjusted model (Covariates: Age, Sex, BMI, Education, Income, Average time spent outdoors, Physical activity minutes, Smoking status, Disability status, Townsend Deprivation Index, and Population Density)
```{r}
summary_confint(lm(Grey_matter_volume_adjusted_T2_cm3~Natural_environment_1000m_10pcnt + Age + Sex + BMI + Education + Income + Average_time_outdoors + Summed_PA_min + Smoking + Disability + Deprivation_Index + Population_density, NoHomeChange_T2))
```

#### Association of natural environment with white matter volume (significant)

Crude model (Covariates: Age, Sex, BMI, Education, and Income)
```{r}
summary_confint(lm(White_matter_volume_adjusted_T2_cm3~Natural_environment_1000m_10pcnt + Age + Sex + BMI + Education + Income, NoHomeChange_T2))
```

Full adjusted model (Covariates: Age, Sex, BMI, Education, Income, Average time spent outdoors, Physical activity minutes, Smoking status, Disability status, Townsend Deprivation Index, and Population Density)
```{r}
summary_confint(lm(White_matter_volume_adjusted_T2_cm3~Natural_environment_1000m_10pcnt + Age + Sex + BMI + Education + Income + Average_time_outdoors + Summed_PA_min + Smoking + Disability + Deprivation_Index + Population_density, NoHomeChange_T2))
```

#### Association of natural environment with hippocampal volume (not significant)

Crude model (Covariates: Age, Sex, BMI, Education, and Income)
```{r}
summary_confint(lm(Avg_Hippocampal_volume_adjusted_T2_cm3~Natural_environment_1000m_10pcnt + Age + Sex + BMI + Education + Income, NoHomeChange_T2))
```

Full adjusted model (Covariates: Age, Sex, BMI, Education, Income, Average time spent outdoors, Physical activity minutes, Smoking status, Disability status, Townsend Deprivation Index, and Population Density)
```{r}
summary_confint(lm(Avg_Hippocampal_volume_adjusted_T2_cm3~Natural_environment_1000m_10pcnt + Age + Sex + BMI + Education + Income + Average_time_outdoors + Summed_PA_min + Smoking + Disability + Deprivation_Index + Population_density, NoHomeChange_T2))
```

#### Moderation by sex: Grey Matter (*not significant, p = 0.054)
```{r}
summary_confint(lm(Grey_matter_volume_adjusted_T2_cm3~Natural_environment_1000m_10pcnt*Sex + Age  + BMI + Education + Income + 
             Average_time_outdoors + 
             Summed_PA_min + Smoking + Disability + Deprivation_Index + Population_density, NoHomeChange_T2))
```

#### Moderation by sex: White Matter (not significant)
```{r}
summary_confint(lm(White_matter_volume_adjusted_T2_cm3~Natural_environment_1000m_10pcnt*Sex + Age  + BMI + Education + Income + 
             Average_time_outdoors + 
             Summed_PA_min + Smoking + Disability + Deprivation_Index + Population_density, NoHomeChange_T2))
```

#### Moderation by sex: Average hippocampal volume (not significant)
```{r}
summary_confint(lm(Avg_Hippocampal_volume_adjusted_T2_cm3~Natural_environment_1000m_10pcnt*Sex + Age  + BMI + Education + Income + 
             Average_time_outdoors + 
             Summed_PA_min + Smoking + Disability + Deprivation_Index + Population_density, NoHomeChange_T2))
```

### 10.2.2) 300m natural environment buffer

#### Association of natural environment with grey matter volume (300m buffer AND restricted sample)

Full adjusted model (Covariates: Age, Sex, BMI, Education, Income, Average time spent outdoors, Physical activity minutes, Smoking status, Disability status, Townsend Deprivation Index, and Population Density)
```{r}
summary_confint(lm(Grey_matter_volume_adjusted_T2_cm3~Natural_environment_300m_10pcnt + Age + Sex + BMI + Education + Income + Average_time_outdoors + Summed_PA_min + Smoking + Disability + Deprivation_Index + Population_density, NoHomeChange_T2))
```

#### Association of natural environment with white matter volume (300m buffer AND restricted sample)

Full adjusted model (Covariates: Age, Sex, BMI, Education, Income, Average time spent outdoors, Physical activity minutes, Smoking status, Disability status, Townsend Deprivation Index, and Population Density)
```{r}
summary_confint(lm(White_matter_volume_adjusted_T2_cm3~Natural_environment_300m_10pcnt + Age + Sex + BMI + Education + Income + Average_time_outdoors + Summed_PA_min + Smoking + Disability + Deprivation_Index + Population_density, NoHomeChange_T2))
```

#### Association of natural environment with average hippocampal volume (300m buffer AND restricted sample)

Full adjusted model (Covariates: Age, Sex, BMI, Education, Income, Average time spent outdoors, Physical activity minutes, Smoking status, Disability status, Townsend Deprivation Index, and Population Density)
```{r}
summary_confint(lm(Avg_Hippocampal_volume_adjusted_T2_cm3~Natural_environment_300m_10pcnt + Age + Sex + BMI + Education + Income + Average_time_outdoors + Summed_PA_min + Smoking + Disability + Deprivation_Index + Population_density, NoHomeChange_T2))
```

#### Moderation by sex: Grey Matter [300m buffer AND subset with no address change] (significant)
```{r}
summary_confint(lm(Grey_matter_volume_adjusted_T2_cm3~Natural_environment_300m_10pcnt*Sex + Age  + BMI + Education + Income + 
             Average_time_outdoors + 
             Summed_PA_min + Smoking + Disability + Deprivation_Index + Population_density, NoHomeChange_T2))
```


#### Moderation by sex: White Matter [300m buffer AND subset with no address change] (not significant)
```{r}
summary_confint(lm(White_matter_volume_adjusted_T2_cm3~Natural_environment_300m_10pcnt*Sex + Age  + BMI + Education + Income + 
             Average_time_outdoors + 
             Summed_PA_min + Smoking + Disability + Deprivation_Index + Population_density, NoHomeChange_T2))
```

#### Moderation by sex: Average hippocampal volume [300m buffer AND subset with no address change] (not significant)
```{r}
summary_confint(lm(Avg_Hippocampal_volume_adjusted_T2_cm3~Natural_environment_300m_10pcnt*Sex + Age  + BMI + Education + Income + 
             Average_time_outdoors + 
             Summed_PA_min + Smoking + Disability + Deprivation_Index + Population_density, NoHomeChange_T2))
```


# 12) Crude models requested by Reviewer 2: using only 3 covariates: age, sex, and ethnic background

## 12.1) 1000m buffer

### 12.1.1) Cognitive Function

#### Association of natural environment with Trail-making Test B-A (not significant)
```{r}
summary_confint(lm(Trails_B_A~Natural_environment_1000m_10pcnt + Age + Sex + Race, CogsBrainCov.df2))
```

#### Association of natural environment with Symbol Digit Substitution (significant)
```{r}
summary_confint(lm(Symbol.Digit~Natural_environment_1000m_10pcnt + Age + Sex + Race, CogsBrainCov.df2))
```

#### Moderation by sex: Trail-making Test B-A (not significant)
```{r}
summary_confint(lm(Trails_B_A~Natural_environment_1000m_10pcnt*Sex + Age + Race, CogsBrainCov.df2))
```

#### Moderation by sex: Symbol Digit Substitution (not significant)
```{r}
summary_confint(lm(Symbol.Digit~Natural_environment_1000m_10pcnt*Sex + Age + Race, CogsBrainCov.df2))
```

### 12.1.2) Brain Volume

#### Association of natural environment with grey matter volume (significant)
```{r}
summary_confint(lm(Grey_matter_volume_adjusted_T2_cm3~Natural_environment_1000m_10pcnt + Age + Sex + Race, CogsBrainCov.df2))
```

#### Association of natural environment with white matter volume (significant)
```{r}
summary_confint(lm(White_matter_volume_adjusted_T2_cm3~Natural_environment_1000m_10pcnt + Age + Sex + Race, CogsBrainCov.df2))
```

#### Association of natural environment with hippocampal volume (not significant)
```{r}
summary_confint(lm(Avg_Hippocampal_volume_adjusted_T2_cm3~Natural_environment_1000m_10pcnt + Age + Sex + Race, CogsBrainCov.df2))
```

#### Moderation by sex: Grey Matter (significant)
```{r}
summary_confint(lm(Grey_matter_volume_adjusted_T2_cm3~Natural_environment_1000m_10pcnt*Sex + Age + Race, CogsBrainCov.df2))
```

#### Moderation by sex: White Matter (not significant)
```{r}
summary_confint(lm(White_matter_volume_adjusted_T2_cm3~Natural_environment_1000m_10pcnt*Sex + Age + Race, CogsBrainCov.df2))
```

#### Moderation by sex: Average hippocampal volume (not significant)
```{r}
summary_confint(lm(Avg_Hippocampal_volume_adjusted_T2_cm3~Natural_environment_1000m_10pcnt*Sex + Age + Race, CogsBrainCov.df2))
```

## 12.2) 300m buffer

### 12.2.1) Cognitive Function

#### Association of natural environment with Trail-making Test B-A (not significant)
```{r}
summary_confint(lm(Trails_B_A~Natural_environment_300m_10pcnt + Age + Sex + Race, CogsBrainCov.df2))
```

#### Association of natural environment with Symbol Digit Substitution (significant)
```{r}
summary_confint(lm(Symbol.Digit~Natural_environment_300m_10pcnt + Age + Sex + Race, CogsBrainCov.df2))
```

#### Moderation by sex: Trail-making Test B-A (not significant)
```{r}
summary_confint(lm(Trails_B_A~Natural_environment_300m_10pcnt*Sex + Age + Race, CogsBrainCov.df2))
```

#### Moderation by sex: Symbol Digit Substitution (not significant)
```{r}
summary_confint(lm(Symbol.Digit~Natural_environment_300m_10pcnt*Sex + Age + Race, CogsBrainCov.df2))
```

### 12.2.2) Brain Volume

#### Association of natural environment with grey matter volume (significant)
```{r}
summary_confint(lm(Grey_matter_volume_adjusted_T2_cm3~Natural_environment_300m_10pcnt + Age + Sex + Race, CogsBrainCov.df2))
```

#### Association of natural environment with white matter volume (significant)
```{r}
summary_confint(lm(White_matter_volume_adjusted_T2_cm3~Natural_environment_300m_10pcnt + Age + Sex + Race, CogsBrainCov.df2))
```

#### Association of natural environment with hippocampal volume (not significant)
```{r}
summary_confint(lm(Avg_Hippocampal_volume_adjusted_T2_cm3~Natural_environment_300m_10pcnt + Age + Sex + Race, CogsBrainCov.df2))
```

#### Moderation by sex: Grey Matter (significant)
```{r}
summary_confint(lm(Grey_matter_volume_adjusted_T2_cm3~Natural_environment_300m_10pcnt*Sex + Age + Race, CogsBrainCov.df2))
```

#### Moderation by sex: White Matter (not significant)
```{r}
summary_confint(lm(White_matter_volume_adjusted_T2_cm3~Natural_environment_300m_10pcnt*Sex + Age + Race, CogsBrainCov.df2))
```

#### Moderation by sex: Average hippocampal volume (not significant)
```{r}
summary_confint(lm(Avg_Hippocampal_volume_adjusted_T2_cm3~Natural_environment_300m_10pcnt*Sex + Age + Race, CogsBrainCov.df2))
```